---
title: "_tables and figures"
author: "Tyler Peckenpaugh"
date: "5/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Additional tables and figures {#etc}

Addtional figures and tables appear here without discussion, for the curious reader.

```{r irtzoomLft, fig.cap="Left tail of raw IRT distribution"}
ggplot(subset(irt_data_all,irt<1000), aes(irt)) +
  geom_histogram(binwidth = 50,color="white",fill="#333333") + 
  xlab("Raw IRT") + ylab("Frequency") + 
  ggtitle("Left tail of IRT distribution (ms)",subtitle="IRT < 1s, bin size = 50ms")
```


```{r zoomRgt, fig.cap="Right tail of raw IRT distribution"}
ggplot(subset(irt_data_all,irt>22000), aes(irt/1000)) +
   geom_histogram(binwidth = 1, color="white", fill="#333333") +
  xlab("Raw IRT") + ylab("Frequency") + 
  ggtitle("Right tail of IRT distribution (s)", subtitle="IRT > 22s, bin size = 1s")
```

## common log
For the purposes of regression analysis, a common log transformation reduces the skew in the data. This distribution is seen in Figure \ref{fig:log10wins}.

```{r log10wins,fig.cap="Common log of wIRT"}
ggplot(irt_data, aes(wirt.log10)) +
  geom_histogram(binwidth = 0.125, color="white",fill="#333333") + 
  xlab("Common log of wIRT") + ylab("Frequency") + 
  ggtitle("Distribution of Common Log of wIRT", subtitle="Bin size = 0.125")
```


# DELCAT

```{r}
cheatdata <- subset(mdata,r1DelCat == "FAST" & Reading == 2)
pp1tab<-xtabs(PP1 ~ Condition_GP + Condition_Q, data=cheatdata)
objtab<-xtabs(OBJ ~ Condition_GP + Condition_Q, data=cheatdata)
tot<-as.data.frame(xtabs( ~ Condition_GP + Condition_Q, data=cheatdata))


pp1<-as.data.frame(pp1tab)
pp1$tot<-tot$Freq
pp1$pct <- pp1$Freq/pp1$tot
pp1$pretty <- paste0(
  pp1$Freq, 
  " (", 
  sprintf("%.1f",pp1$pct * 100), 
  "%)"
)
pp1disp <- pp1 %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

obj<-as.data.frame(objtab)
obj$tot<-tot$Freq
obj$pct <- obj$Freq/obj$tot
obj$pretty <- paste0(
  obj$Freq, 
  " (", 
  sprintf("%.1f",obj$pct * 100), 
  "%)"
)
objdisp <- obj %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

pp1obj <- cbind(pp1disp,objdisp[2:3])
pp1obj[[1]] <- c("+GP","-GP")
fast <- pp1obj
```

```{r}
cheatdata <- subset(mdata,r1DelCat == "SLOW" & Reading == 2)
pp1tab<-xtabs(PP1 ~ Condition_GP + Condition_Q, data=cheatdata)
objtab<-xtabs(OBJ ~ Condition_GP + Condition_Q, data=cheatdata)
tot<-as.data.frame(xtabs( ~ Condition_GP + Condition_Q, data=cheatdata))


pp1<-as.data.frame(pp1tab)
pp1$tot<-tot$Freq
pp1$pct <- pp1$Freq/pp1$tot
pp1$pretty <- paste0(
  pp1$Freq, 
  " (", 
  sprintf("%.1f",pp1$pct * 100), 
  "%)"
)
pp1disp <- pp1 %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

obj<-as.data.frame(objtab)
obj$tot<-tot$Freq
obj$pct <- obj$Freq/obj$tot
obj$pretty <- paste0(
  obj$Freq, 
  " (", 
  sprintf("%.1f",obj$pct * 100), 
  "%)"
)
objdisp <- obj %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

pp1obj <- cbind(pp1disp,objdisp[2:3])
pp1obj[[1]] <- c("+GP","-GP")
```

The break patterns for slow and fast participants are shown in table \ref{tab:cheaterBreaks}.

```{r cheaterBreaks}
cat <-cbind(fast, pp1obj[2:5])
kable(
  cat,
  caption="Percentage of R2 recordings with PP1/OBJ break by condition and R1 delay category",
  col.names = c("", "D", "Q","D", "Q","D", "Q","D", "Q")
) %>% 
  add_header_above(c(" "=1,"PP1 Break"=2,"OBJ Break"=2,"PP1 Break"=2,"OBJ Break"=2)) %>%
  add_header_above(c(" "=1,"Fast R1 delay"=4,"Slow R1 delay"=4)) %>%
  kable_styling(latex_options = c("hold_position","scale_down"))
```

```{r}
cheatdata <- subset(mdata,r1DelCat == "FAST" & Reading == 2)
twolv<-xtabs( ~ two_level_prosody + condition, data=cheatdata) %>%
  prop.table(margin=2)

twolvFAST <- twolv * 100

cheatdata <- subset(mdata,r1DelCat == "SLOW" & Reading == 2)
twolv<-xtabs( ~ two_level_prosody + condition, data=cheatdata) %>%
  prop.table(margin=2)

twolv <- cbind(twolvFAST, twolv * 100)

twolv %>% 
  kable(
    caption="Prosodic pattern by condition, SLOW/FAST R1 delay category (Reading 2)",
    digits = 2
  ) %>%
  add_header_above(c(" " = 1, "FAST" = 4, "SLOW" = 4)) %>%
  kable_styling(latex_options = c("hold_position"))
```
```{r}
cheatdata <- subset(mdata,r1DelCat == "FAST" & Reading == 1)
twolv<-xtabs( ~ two_level_prosody + condition, data=cheatdata) %>%
  prop.table(margin=2)

twolvFAST <- twolv * 100

cheatdata <- subset(mdata,r1DelCat == "SLOW" & Reading == 1)
twolv<-xtabs( ~ two_level_prosody + condition, data=cheatdata) %>%
  prop.table(margin=2)

twolv <- cbind(twolvFAST, twolv * 100)

twolv %>% 
  kable(
    caption="Prosodic pattern by condition, SLOW/FAST R1 delay category (Reading 1)",
    digits = 2
  ) %>%
  add_header_above(c(" " = 1, "FAST" = 4, "SLOW" = 4)) %>%
  kable_styling(latex_options = c("hold_position"))
```

```{r r1delByP}
df <- r1delByP
df$r1delbyp <- df$r1delbyp / 1000

df %>%  
  arrange(df$r1delbyp) %>% 
  kable(
    col.names = c("Participant", "Mean R1 delay","R1 delay category"), 
    caption="Mean reading 1 delay (seconds) and R1 delay category by participant", 
    digits = 2
  ) %>% kable_styling(latex_options = c("repeat_header"))%>% 
  footnote(catdes)
```