---
title: "Summary"
author: "Tyler Peckenpaugh"
date: "3/5/2019"
output:
  pdf_document: default
  html_document:
    highlight: tango
    theme: journal
---

```{r setup, include=FALSE}
library(readr)
library(knitr)
library(pander)
library(lme4)
library(lmerTest)
library(lattice)

got_prosody <- read_csv("dissertation/csvs/got_prosody.csv")

mdata <- read_csv("dissertation/csvs/got_prosody.csv")
mdata$V<-mdata$V == "YES"
mdata$OBJ<-mdata$OBJ == "YES"
mdata$PP1<-mdata$PP1 == "YES"
mdata$STRUG<-mdata$STRUG == "YES"
mdata$STRONG <- as.factor(mdata$STRONG)
mdata$WEAK <- as.factor(mdata$WEAK)
mdata$Reading <- as.factor(mdata$Reading)
mdata$prosody <- as.factor(mdata$prosody)
mdata$two_level_prosody <- as.factor(mdata$two_level_prosody)
mdata$IID <- as.factor(mdata$IID)
mdata$UID <- as.factor(mdata$UID)
mdata$SID <- as.factor(mdata$SID)
mdata$STRUG_START <- as.factor(mdata$STRUG_START)
mdata$goodProsody <- mdata$two_level_prosody %in% c("PP1", "PP1 > OBJ")

inter_item_timings_all <- read_csv("dissertation/csvs/inter_item_timings.csv", 
    col_types = cols(item = col_character(), 
        participant = col_character()))

inter_item_timings_all$inter_reading[inter_item_timings_all$inter_reading <0] <- NA
inter_item_timings_all$Exp <- FALSE == grepl("F",inter_item_timings_all$filename)
inter_item_timings <- subset(inter_item_timings_all, Exp & participant != "108")

gpInsensitive <- paste(c(1,
12,
14,
15,
19,
2,
20,
203,
21,
210,
22,
4,
6))

```

# 3-level prosodic pattern
```{r 3lvlxtabs}
prosodies.full <- xtabs(~ prosody + Condition_Q + Condition_GP, mdata)
prosodies.r1 <- xtabs(~ prosody + Condition_Q + Condition_GP, subset(mdata,Reading==1))
prosodies.r2 <- xtabs(~ prosody + Condition_Q + Condition_GP, subset(mdata,Reading==2))
pander(prosodies.full, caption = "both readings")
pander(prosodies.r1, caption = "reading 1")
pander(prosodies.r2, caption = "reading 2")
```

# 2-level prosodic pattern
```{r 2lvlxtabs}
prosodies.full <- xtabs(~ two_level_prosody + Condition_Q + Condition_GP, mdata)
prosodies.r1 <- xtabs(~ two_level_prosody + Condition_Q + Condition_GP, subset(mdata,Reading==1))
prosodies.r2 <- xtabs(~ two_level_prosody + Condition_Q + Condition_GP, subset(mdata,Reading==2))
pander(prosodies.full)
pander(prosodies.r1, caption = "reading 1")
pander(prosodies.r2, caption = "reading 2")
```


# Timing models (ME Linear Regression)
```{r lmerModel,echo=FALSE}
if(!exists("models.timing")){
  models.timing <- lmer(inter_reading ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(!exists("models.timingNoInt")){
  models.timingNoInt <- lmer(inter_reading ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}

if(!exists("models.timing.Sense")){
  models.timing.Sense <- lmer(inter_reading ~ GP_condition + Q_condition + (1 + GP_condition + Q_condition | participant) + (1 + GP_condition + Q_condition | item), data=subset(inter_item_timings, !(participant %in% gpInsensitive)), control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(!exists("models.timingNoInt.Sense")){
  models.timingNoInt.Sense <- lmer(inter_reading ~ GP_condition + Q_condition + (1 + GP_condition + Q_condition | participant) + (1 + GP_condition + Q_condition | item), data=subset(inter_item_timings, !(participant %in% gpInsensitive)), control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
```
```{r assumptionChecks}
inter_item_timings$resids_main[!is.na(inter_item_timings$inter_reading)] <- residuals(models.timing)
inter_item_timings$resids_main.abs <- abs(inter_item_timings$resids_main)
inter_item_timings$resids_main.sq <-inter_item_timings$resids_main.abs^2
sensitive_timings <- subset(inter_item_timings, !(participant %in% gpInsensitive))

timings_sq_resid <- lm(inter_item_timings$resids_main.sq ~ inter_item_timings$participant, data=inter_item_timings) #ANOVA of the squared residuals
sens_timings_sq_resid <- lm(resids_main.sq ~ participant, data=sensitive_timings) #ANOVA of the squared residuals
```
# Homogeneity of variance
Anova of squared residuals ~ participant for full data and "GP-sensitive" subset
```{r homovar}
anova(timings_sq_resid) #displays the results
anova(sens_timings_sq_resid) #displays the results
```
# Normal distribution of residuals
All participants
`r qqmath(models.timing, id=0.05)`
Only GP-sensitive participants
`r qqmath(models.timing.Sense, id=0.05)`
```
# Models
## Full model
```{r full}
summary(models.timing)
```
## Without interacton term
```{r noInt}
summary(models.timingNoInt)
```
## Only "gp-sensitive"
The following excludes all participants whose mean inter-item reading time for `+GP` items is less than for `-GP` items
### With interaction
```{r gpSense}
summary(models.timing.Sense)
```
## Without interaction
```{r gpSenseNoInt}
summary(models.timingNoInt.Sense)

```
# Logit models
```{r logitModels}
# models.full <- glmer(PP1~V*OBJ+Condition_Q*Condition_GP+Reading + (1+V*OBJ+Condition_Q*Condition_GP+Reading | SID) + (1+V*OBJ+Condition_Q*Condition_GP+Reading | IID),data=mdata,family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=2e8)))
# summary(models.full)
if(!exists("models.goodP")){
  models.goodP <- glmer(goodProsody~Condition_Q*Condition_GP*Reading + (1+Condition_Q*Condition_GP*Reading | SID) + (1+Condition_Q*Condition_GP*Reading | IID),data=mdata,family = binomial, control = glmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=2e8)))
}
summary(models.goodP)
```