---
title: "Summary"
author: "Tyler Peckenpaugh"
date: "3/5/2019"
output:
  html_document:
    theme: journal
---

```{r setup, include=FALSE}
library(readr)
library(knitr)
library(kableExtra)
library(pander)
library(lme4)
library(lmerTest)
library(lattice)

refresh = FALSE # set to true to re-run models

mdata <- read_csv("dissertation/csvs/got_prosody.csv")
mdata$V<-mdata$V == "YES"
mdata$OBJ<-mdata$OBJ == "YES"
mdata$PP1<-mdata$PP1 == "YES"
mdata$STRUG<-mdata$STRUG == "YES"
mdata$STRONG <- as.factor(mdata$STRONG)
mdata$WEAK <- as.factor(mdata$WEAK)
mdata$Reading <- as.factor(mdata$Reading)
mdata$prosody <- as.factor(mdata$prosody)
mdata$two_level_prosody <- as.factor(mdata$two_level_prosody)
mdata$IID <- as.factor(mdata$IID)
mdata$UID <- as.factor(mdata$UID)
mdata$SID <- as.factor(mdata$SID)
mdata$STRUG_START <- as.factor(mdata$STRUG_START)
mdata$lowAttach <- mdata$two_level_prosody %in% c("PP1", "PP1 > OBJ")
mdata$IID[mdata$IID == 108] <- 8
mdata$LQI <- mdata$QI == "YES"
mdata$HtOrMs_QI <- mdata$LQI == mdata$Condition_Q

mdata$condition <- paste(ifelse(mdata$Condition_GP, "+GP", "-GP")," ",ifelse(mdata$Condition_Q, "+Q", "-Q"))

gpInsensitive <- paste(c(1,
12,
14,
15,
19,
2,
20,
203,
21,
210,
22,
4,
6))

```

# Crosstabs

## Struggle 

Note that struggle does not show good cross-rater reliability

```{r strug}
label = "struggle:"
f <- ~STRUG + condition
full <- xtabs(f, mdata)
r1 <- xtabs(f, subset(mdata,Reading==1))
r2 <- xtabs(f, subset(mdata,Reading==2))
kable(full, caption = paste(label,"both readings")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r1, caption = paste(label,"reading 1")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r2, caption = paste(label,"reading 2")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
```

## V break

Note that V break does not show good cross-rater reliability

```{r v}
label = "pp1 break:"
f <- ~V + condition
full <- xtabs(f, mdata)
r1 <- xtabs(f, subset(mdata,Reading==1))
r2 <- xtabs(f, subset(mdata,Reading==2))
kable(full, caption = paste(label,"both readings")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r1, caption = paste(label,"reading 1")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r2, caption = paste(label,"reading 2")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
```

## PP1 break

```{r pp1}
label = "pp1 break:"
f <- ~PP1 + condition
full <- xtabs(f, mdata)
r1 <- xtabs(f, subset(mdata,Reading==1))
r2 <- xtabs(f, subset(mdata,Reading==2))
kable(full, caption = paste(label,"both readings")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r1, caption = paste(label,"reading 1")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r2, caption = paste(label,"reading 2")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
```
## OBJ break

```{obj}
label <- "obj:"
f <- ~OBJ + condition
full <- xtabs(f, mdata)
r1 <- xtabs(f, subset(mdata,Reading==1))
r2 <- xtabs(f, subset(mdata,Reading==2))
kable(full, caption =paste(label,"both readings")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r1, caption = paste(label,"reading 1")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r2, caption = paste(label,"reading 2")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
```

## Question intonation hit/miss

```{r qihit}
label <- "QI hit:"
f <- ~HtOrMs_QI + condition
full <- xtabs(f, mdata)
r1 <- xtabs(f, subset(mdata,Reading==1))
r2 <- xtabs(f, subset(mdata,Reading==2))
kable(full, caption = paste(label,"both readings")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r1, caption = paste(label,"reading 1")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(r2, caption = paste(label,"reading 2")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
```

# 3-level prosodic pattern
```{r 3lvlxtabs}
f <- ~ prosody + condition
prosodies.full <- xtabs(f, mdata)
prosodies.r1 <- xtabs(f, subset(mdata,Reading==1))
prosodies.r2 <- xtabs(f, subset(mdata,Reading==2))
kable(prosodies.full, caption = "both readings") %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(prosodies.r1, caption = "reading 1") %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(prosodies.r2, caption = "reading 2") %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
```

# 2-level prosodic pattern
```{r 2lvlxtabs}
form <- ~ condition+ two_level_prosody
prosodies.full <- xtabs(form, mdata)
prosodies.r1 <- xtabs(form, subset(mdata,Reading==1))
prosodies.r2 <- xtabs(form, subset(mdata,Reading==2))
kable(prosodies.full) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(prosodies.r1, caption = "reading 1") %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
kable(prosodies.r2, caption = "reading 2") %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T) 
```

# Logit models

```{r pp1logit}
if(!exists("models.logi.pp1.full") | refresh){
  models.logi.pp1.full <- glmer(
    PP1~OBJ * Condition_Q * Condition_GP * Reading + 
      (1+OBJ*Condition_Q*Condition_GP*Reading | SID) + 
      (1+OBJ*Condition_Q*Condition_GP*Reading | IID),
    family = binomial, 
    data = mdata,
    control = glmerControl(optimizer = "bobyqa", calc.derivs=F,optCtrl=list(maxfun=2e8))
  )
}
```
```{r attachment}
if(!exists("models.logi.attach.full") | refresh){
  models.logi.attach.full <- glmer(
    lowAttach~Condition_Q*Condition_GP*Reading + 
      (1+Condition_Q*Condition_GP*Reading | SID) + 
      (1+Condition_Q*Condition_GP*Reading | IID),
    data = mdata,
    family = binomial, 
    control = glmerControl(optimizer = "bobyqa", calc.derivs=F, optCtrl=list(maxfun=2e8))
  )
}
```

## Predicting PP1

Is the existence of PP1 break predicted by condition * reading?

```{r summarizePP1}
summary(models.logi.pp1.full)
```

## Predicting PP1 as strongest or only break

Is PP1 being the strongest or only break predicted by condition * reading?

```{r summarizeAttach}
summary(models.logi.attach.full)
```