---
title: "LME Inter-Item Timing (winsorized)"
author: "Tyler Peckenpaugh"
date: "3/5/2019"
output:
  html_document:
    theme: journal
    toc: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(readr)
library(knitr)
library(kableExtra)
library(pander)
library(lme4)
library(lmerTest)
library(lattice)
library(psych)


mean.sd <- function(x) c(mean = mean(x), sd = sd(x))
# set to true to re-run modelsqqnorm(myModel.lme, ~ranef(., level=2))
refresh = TRUE
inter_item_timings_all <- read_csv("csvs/irt-20190318-151704.csv")


inter_item_timings_all$condition <- paste(ifelse(inter_item_timings_all$is_filler,ifelse(inter_item_timings_all$GP_condition,"+PP","-PP"),ifelse(inter_item_timings_all$GP_condition,"+GP","-GP")),ifelse(inter_item_timings_all$Q_condition,"+Q","-Q"))

inter_item_timings_all$inter_reading[inter_item_timings_all$inter_reading < 0] <- NA
inter_item_timings_all$inter_reading[inter_item_timings_all$inter_reading < 1] <- 1
inter_item_timings_all$Exp <- FALSE == grepl("F",inter_item_timings_all$filename)
inter_item_timings_all$participant[inter_item_timings_all$participant == 108] <- 8

inter_item_timings_all<-transform(inter_item_timings_all,winsrdIRT = ave(inter_reading,participant,FUN=winsor))

inter_item_timings_all$raw_irt <- inter_item_timings_all$inter_reading
inter_item_timings_all$inter_reading <- inter_item_timings_all$winsrdIRT

inter_item_timings <- subset(inter_item_timings_all, Exp)

inter_item_timings$inter_reading_shift <- 1 + inter_item_timings$inter_reading 
inter_item_timings$inter_reading.ln <- log(inter_item_timings$inter_reading_shift)
inter_item_timings$inter_reading.log10 <- log10(inter_item_timings$inter_reading_shift)
inter_item_timings$inter_reading.inverse <- 1/(inter_item_timings$inter_reading_shift)
inter_item_timings$inter_reading.sqred <- (inter_item_timings$inter_reading_shift)^2

gpInsensitive <- paste(c(1,12,14,15,19,2,20,203,21,210,22,4,6))

```



# Inter-item timing analyses

Mean inter-reading time (IRT) is a measure of the amount of time in ms between when a subject stops speaking during a cold reading and when they begin speaking for a previewed reading.

Subjects were asked to read each sentence twice, once with no preview at all, and then again after unlimited preview. IRT is meant to represent their chosen preview time for the second reading.

## Mean and SD of IRT by condition

Mean inter-reading time (IRT) in ms

```{r meantable, echo=FALSE}
mtable <- with(na.omit(inter_item_timings),tapply(inter_reading, condition, mean))

kable(mtable, align="r",col.names = c("mean")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
```

Standard deviation inter-reading time (IRT) in ms

```{r sdtable, echo=FALSE}
mtable <- with(na.omit(inter_item_timings),tapply(inter_reading, list("+GP"=GP_condition,"+Q"=Q_condition), sd))
row.names(mtable) <- c("-GP","+GP")
kable(mtable, align="r",col.names = c("-Q","+Q")) %>%
  kable_styling(bootstrap_options = c("striped",full_width=F,position="left")) %>%
  column_spec(1,bold=T)
```

## boxplot of mean inter-item time by condition

Inter reading time (IRT) in ms

```{r echo=FALSE}
boxplot(inter_reading ~condition, main="plain IRT", col=c("white","lightgray"),inter_item_timings)

boxplot(inter_reading.ln ~ condition, main="natural log of IRT", col=c("white","lightgray"),inter_item_timings)
boxplot(inter_reading.inverse ~ condition,main="inverse of IRT", col=c("white","lightgray"),inter_item_timings)
boxplot(inter_reading.log10 ~ condition,main="common log of IRT", col=c("white","lightgray"),inter_item_timings)
boxplot(inter_reading.sqred ~ condition,main="IRT squared", col=c("white","lightgray"),inter_item_timings)

```
# Models

```{r lmerModel,echo=FALSE,include=FALSE}
if(refresh | !exists("models.timing")){
  models.timing <- lmer(inter_reading ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(refresh | !exists("models.timingNoInt")){
  models.timingNoInt <- lmer(inter_reading ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}

if(refresh | !exists("models.timing.Sense")){
  models.timing.Sense <- lmer(inter_reading ~ GP_condition + Q_condition + (1 + GP_condition + Q_condition | participant) + (1 + GP_condition + Q_condition | item), data=subset(inter_item_timings, !(participant %in% gpInsensitive)), control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(refresh | !exists("models.timingNoInt.Sense")){
  models.timingNoInt.Sense <- lmer(inter_reading ~ GP_condition + Q_condition + (1 + GP_condition + Q_condition | participant) + (1 + GP_condition + Q_condition | item), data=subset(inter_item_timings, !(participant %in% gpInsensitive)), control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(refresh | !exists("models.timingLn")){
  models.timingLn <- lmer(inter_reading.ln ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(refresh | !exists("models.timingInv")){
  models.timingInv <- lmer(inter_reading.inverse ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(refresh | !exists("models.timingLg10")){
  models.timingLg10 <- lmer(inter_reading.log10 ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(refresh | !exists("models.timingSq")){
  models.timingSq <- lmer(inter_reading.sqred ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
```
```{r assumptionChecks, include=FALSE, echo=FALSE}
# residuals
inter_item_timings$resids_main <- NA
inter_item_timings$resids_main[!is.na(inter_item_timings$inter_reading)] <- residuals(models.timing)
inter_item_timings$resids_main.abs <- abs(inter_item_timings$resids_main)
inter_item_timings$resids_main.sq <-inter_item_timings$resids_main.abs^2

# residuals (squared)
inter_item_timings$resids_mainSq <- NA
inter_item_timings$resids_mainSq[!is.na(inter_item_timings$inter_reading.sqred)] <- residuals(models.timingSq)
inter_item_timings$resids_mainSq.abs <- abs(inter_item_timings$resids_mainSq)
inter_item_timings$resids_mainSq.sq <-inter_item_timings$resids_mainSq.abs^2

# residuals (ln)
inter_item_timings$resids_mainLn <- NA
inter_item_timings$resids_mainLn[!is.na(inter_item_timings$inter_reading.ln)] <- residuals(models.timingLn)
inter_item_timings$resids_mainLn.abs <- abs(inter_item_timings$resids_mainLn)
inter_item_timings$resids_mainLn.sq <-inter_item_timings$resids_mainLn.abs^2

# residuals (inverse)
inter_item_timings$resids_mainInv <- NA
inter_item_timings$resids_mainInv[!is.na(inter_item_timings$inter_reading.inverse)] <- residuals(models.timingInv)
inter_item_timings$resids_mainInv.abs <- abs(inter_item_timings$resids_mainInv)
inter_item_timings$resids_mainInv.sq <-inter_item_timings$resids_mainInv.abs^2

# residuals (log10)
inter_item_timings$resids_mainLg10 <- NA
inter_item_timings$resids_mainLg10[!is.na(inter_item_timings$inter_reading.inverse)] <- residuals(models.timingLg10)
inter_item_timings$resids_mainLg10.abs <- abs(inter_item_timings$resids_mainLg10)
inter_item_timings$resids_mainLg10.sq <-inter_item_timings$resids_mainLg10.abs^2

# gp sense
sensitive_timings <- subset(inter_item_timings, !(participant %in% gpInsensitive))
sensitive_timings$resids_main[!is.na(sensitive_timings$inter_reading)] <- residuals(models.timing.Sense)
sensitive_timings$resids_main.abs <- abs(sensitive_timings$resids_main)
sensitive_timings$resids_main.sq <- sensitive_timings$resids_main.abs^2

# ANOVA of the squared residuals
timings_sq_resid <- lm(resids_main.sq ~ participant, data=inter_item_timings) 
# ANOVA of the squared residuals (Log10)
timings_sq_resid_log10 <- lm(resids_mainLg10.sq ~ participant, data=inter_item_timings) 
# ANOVA of the squared residuals (squared)
timings_sq_resid_sq <- lm(resids_mainSq.sq ~ participant, data=inter_item_timings) 
# ANOVA of the squared residuals (inverse)
timings_sq_resid_inv <- lm(resids_mainInv.sq ~ participant, data=inter_item_timings) 
# ANOVA of the squared residuals (ln)
timings_sq_resid_ln <- lm(resids_mainLn.sq ~ participant, data=inter_item_timings) 
# ANOVA of the squared residuals (gp-sensitive)
sens_timings_sq_resid <- lm(resids_main.sq ~ participant, data=sensitive_timings)
```



## All participants

These models use all 32 participants

### Full model

```{r full}
summary(models.timing)
```

### Full model (natural log)

```{r fullln}
summary(models.timingLn)
```

### Full model (common log)

```{r fulll10}
summary(models.timingLg10)
```

### Full model (inverse)

```{r fullinv}
summary(models.timingInv)
```

### Full model (squared)

```{r fullsq}
summary(models.timingSq)
```

### Without interacton 

```{r noInt}
summary(models.timingNoInt)
```

## Only "gp-sensitive"
The following models exclude all participants whose mean inter-item reading time for `+GP` items is less than for `-GP` items

### With interaction
```{r gpSense}
summary(models.timing.Sense)
```

### Without interaction
```{r gpSenseNoInt}
summary(models.timingNoInt.Sense)

```

# Checking assumptions

The following checks assumptions about the data

## Homogeneity of variance

Anova of squared residuals (full model) by participant. If the data fit the assumption, we expect no statistical significance.

### All Ps

```{r homovar}
# all participants
anova(timings_sq_resid) 
```

### All Ps (natural log)

```{r homovarLn}
# all participants
anova(timings_sq_resid_ln) 
```

### All Ps (squared)

```{r homovarSq}
# all participants
anova(timings_sq_resid_sq) 
```

### All Ps (inverse)

```{r homovarInv}
# all participants
anova(timings_sq_resid_inv) 
```

### All Ps (Log10)

```{r homovarLg10}
# all participants
anova(timings_sq_resid_log10) 
```

### Only GP-sense

```{r homovarSense}
# GP-sensitive only
anova(sens_timings_sq_resid)
```

## Normal distribution of residuals

QQ plots of models with interation

### All participants

Outcome is plain IRT
```{r normdres}
qqmath(models.timing, id=0.05)
```

### All participants (Natural log)

Outcome is plain natural log of IRT
```{r normdresLn}
qqmath(models.timingLn, id=0.05)
```

### All participants (Log 10)

Outcome is common log of IRT
```{r normdresLg10}
qqmath(models.timingLg10, id=0.05)
```

### All participants (Inverse)

Outcome is inverse of IRT (i.e. 1/IRT)
```{r normdresInverse}
qqmath(models.timingInv, id=0.05)
```

### All participants (Squared)

Outcome is IRT squared
```{r normdresSq}
qqmath(models.timingSq, id=0.05)
```

### Only GP-sensitive participants

Outcome is plain IRT, but uses subset of participants sensitive to GP
```{r normdresSense}
qqmath(models.timing.Sense, id=0.05)
```

## Linearity

Plots of residuals vs. observed

### All Ps

Plain IRT
```{r lin}
plot(resid(models.timing),na.omit(inter_item_timings$inter_reading),xlab="Residuals",ylab="Observed values")
```

### All Ps (nautral log)

Natural log of IRT
```{r linLn}
plot(resid(models.timingLn),na.omit(inter_item_timings$inter_reading.ln),xlab="Residuals",ylab="Observed values")
```

### All Ps (log10)

Common log of IRT
```{r linLg10}
plot(resid(models.timingLg10),na.omit(inter_item_timings$inter_reading.log10),xlab="Residuals",ylab="Observed values")
```

### All Ps (squared)

IRT squared
```{r linSq}
plot(resid(models.timingSq),na.omit(inter_item_timings$inter_reading.sqred),xlab="Residuals",ylab="Observed values")
```

### All Ps (Inverse)

1/IRT
```{r linInv}
plot(resid(models.timingInv),na.omit(inter_item_timings$inter_reading.inverse),xlab="Residuals",ylab="Observed values")
```

### GP-sensitive only

Subset of subjects sensitive to GP
```{r linSense}
plot(resid(models.timing.Sense),na.omit(sensitive_timings$inter_reading),xlab="Residuals",ylab="Observed values")
```