---
title: "LME Inter-Item Timing, Inverse transform"
author: "Tyler Peckenpaugh"
date: "3/5/2019"
output:
  html_document:
    theme: journal
    toc: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(readr)
library(knitr)
library(pander)
library(lme4)
library(lmerTest)
library(lattice)

# set to true to re-run modelsqqnorm(myModel.lme, ~ranef(., level=2))
refresh = TRUE

inter_item_timings_all <- read_csv("dissertation/csvs/inter_item_timings.csv", 
    col_types = cols(item = col_character(), 
        participant = col_character()))

inter_item_timings_all$inter_reading[inter_item_timings_all$inter_reading < 0] <- NA
inter_item_timings_all$inter_reading <- inter_item_timings_all$inter_reading + 1
inter_item_timings_all$Exp <- FALSE == grepl("F",inter_item_timings_all$filename)
inter_item_timings_all$participant[inter_item_timings_all$participant == 108] <- 8
inter_item_timings <- subset(inter_item_timings_all, Exp)

inter_item_timings$inter_reading_raw <- inter_item_timings$inter_reading
inter_item_timings$inter_reading <- 1/(inter_item_timings$inter_reading)
  
gpInsensitive <- paste(c(1,
12,
14,
15,
19,
2,
20,
203,
21,
210,
22,
4,
6))

```



# Inter-item timing analyses

## boxplot of mean inter-item time by condition
```{r echo=FALSE}
boxplot(inter_reading ~ GP_condition*Q_condition, col=c("white","lightgray"),inter_item_timings,names=c("-GP -Q", "+GP -Q", "-GP +Q", "+GP +Q"))
```
```{r lmerModel,echo=FALSE}
if(refresh | !exists("models.timing")){
  models.timing <- lmer(inter_reading ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(refresh | !exists("models.timingNoInt")){
  models.timingNoInt <- lmer(inter_reading ~ GP_condition * Q_condition + (1 + GP_condition * Q_condition | participant) + (1 + GP_condition * Q_condition | item), data=inter_item_timings, control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}

if(refresh | !exists("models.timing.Sense")){
  models.timing.Sense <- lmer(inter_reading ~ GP_condition + Q_condition + (1 + GP_condition + Q_condition | participant) + (1 + GP_condition + Q_condition | item), data=subset(inter_item_timings, !(participant %in% gpInsensitive)), control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
if(refresh | !exists("models.timingNoInt.Sense")){
  models.timingNoInt.Sense <- lmer(inter_reading ~ GP_condition + Q_condition + (1 + GP_condition + Q_condition | participant) + (1 + GP_condition + Q_condition | item), data=subset(inter_item_timings, !(participant %in% gpInsensitive)), control=lmerControl(optCtrl=list(maxfun=2e12) ), REML=FALSE)
}
```
```{r assumptionChecks, include=FALSE, echo=FALSE}
inter_item_timings$resids_main <- NA
inter_item_timings$resids_main[!is.na(inter_item_timings$inter_reading)] <- residuals(models.timing)
inter_item_timings$resids_main.abs <- abs(inter_item_timings$resids_main)
inter_item_timings$resids_main.sq <-inter_item_timings$resids_main.abs^2
sensitive_timings <- subset(inter_item_timings, !(participant %in% gpInsensitive))
sensitive_timings$resids_main[!is.na(sensitive_timings$inter_reading)] <- residuals(models.timing.Sense)
sensitive_timings$resids_main.abs <- abs(sensitive_timings$resids_main)
sensitive_timings$resids_main.sq <- sensitive_timings$resids_main.abs^2

# ANOVA of the squared residuals
timings_sq_resid <- lm(resids_main.sq ~ participant, data=inter_item_timings) 
# ANOVA of the squared residuals
sens_timings_sq_resid <- lm(resids_main.sq ~ participant, data=sensitive_timings)
```

# Models

## All participants
These models use all 32 participants

### Full model
```{r full}
summary(models.timing)
```

### Without interacton 
```{r noInt}
summary(models.timingNoInt)
```

## Only "gp-sensitive"
The following models exclude all participants whose mean inter-item reading time for `+GP` items is less than for `-GP` items

### With interaction
```{r gpSense}
summary(models.timing.Sense)
```

### Without interaction
```{r gpSenseNoInt}
summary(models.timingNoInt.Sense)

```
# Checking assumptions
The following checks assumptions about the data

## Homogeneity of variance
Anova of squared residuals (full model) by participant 

### All Ps
```{r homovar}
# all participants
anova(timings_sq_resid) 
```

### Only GP-sense
```{r homovarSense}
# GP-sensitive only
anova(sens_timings_sq_resid)
```

## Normal distribution of residuals
QQ plots of full models

### All participants
```{r normdres}
qqmath(models.timing, id=0.05)
```

### Only GP-sensitive participants
```{r normdresSense}
qqmath(models.timing.Sense, id=0.05)
```

## Linearity
Plots of residuals vs. observed

### All Ps
```{r lin}
plot(resid(models.timing),na.omit(inter_item_timings$inter_reading),xlab="Residuals",ylab="Observed values")
```

### GP-sensitive only
```{r linsense}
plot(resid(models.timing.Sense),na.omit(sensitive_timings$inter_reading),xlab="Residuals",ylab="Observed values")
```