---
title: "Prosodic properties"
author: "Tyler Peckenpaugh"
date: "4/2/2019"
output:
output:
  tufte::tufte_html: 
    tufte_features: ["fonts", "italics"]
    toc: True
  tufte::tufte_handout: 
    toc: True  
    latex_engine: xelatex
---

```{r setup, include=FALSE}
library(readr)
library(knitr)
library(kableExtra)
library(lme4)
library(lmerTest)

refresh = FALSE # set to true to re-run models
order1<-paste(c(1,2,3,4,5,6,7,8,9,22,201,202,203,204,205,206))
gpInsensitive <- paste(c(1,12,14,15,19,2,20,203,21,210,22,4,6))

morder<-c(3,8,4,9,14,13,12,2,7,10,1,16,5,15,11,6)
sorder<-c(7,10,1,16,5,15,11,6,3,8,4,9,14,13,12,2)

getIOrd<-function (x,s){
  if(s==1){
    return (match(x,morder))
  } else {
    return (match(x,sorder))
  }
}

mdata <- read_csv("csvs/got_prosody.csv")

mdata$Sorder <- 2
mdata$Sorder[mdata$SID %in% order1] <- 1
mdata$Iorder <- getIOrd(mdata$IID,mdata$Sorder)

mdata<-subset(mdata,!is.na(PP1) & !is.na(two_level_prosody))
mdata$V<-mdata$V == "YES"
mdata$OBJ<-mdata$OBJ == "YES"
mdata$PP1<-mdata$PP1 == "YES"
mdata$STRUG<-mdata$STRUG == "YES"
mdata$STRONG <- as.factor(mdata$STRONG)
mdata$WEAK <- as.factor(mdata$WEAK)
mdata$Reading <- as.factor(mdata$Reading)
mdata$prosody <- as.factor(mdata$prosody)
mdata$two_level_prosody <- as.factor(mdata$two_level_prosody)
mdata$IID <- as.factor(mdata$IID)
mdata$UID <- as.factor(mdata$UID)
mdata$SID <- as.factor(mdata$SID)
mdata$STRUG_START <- as.factor(mdata$STRUG_START)
mdata$lowAttach <- mdata$two_level_prosody %in% c("PP1", "PP1 > OBJ")
mdata$IID[mdata$IID == 108] <- 8
mdata$LQI <- mdata$QI == "YES"
mdata$HtOrMs_QI <- mdata$LQI == mdata$Condition_Q

mdata$condition <- paste(ifelse(mdata$Condition_GP, "+GP", "-GP")," ",ifelse(mdata$Condition_Q, "+Q", "-Q"))




```

# Logit models

```{r pp1logit}
# model failse to converge
# if(F & !exists("models.logi.attach.full") | refresh){
#   models.logi.pp1.full <- glmer(
#     PP1~OBJ + Condition_Q * Condition_GP + Reading + 
#       (OBJ+Condition_Q*Condition_GP+Reading | SID) + 
#       (OBJ+Condition_Q*Condition_GP+Reading | IID),
#     family = binomial, 
#     data = mdata,
#     control = glmerControl(optimizer = "bobyqa",
#                            optCtrl=list(maxfun=2e8)),
#     verbose = 2
#   )
# }
```
```{r attachment,echo=FALSE}
if(!exists("models.logi.attach.full.wD") | refresh){
  models.logi.attach.full.wD <- glmer(
    lowAttach~Condition_Q*Condition_GP*Reading + 
      (1+Condition_Q*Condition_GP*Reading | SID) + 
      (1+Condition_Q*Condition_GP*Reading | IID),
    data = mdata,
    family = binomial, 
    control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e8)),
    verbose = 2
  )
}
```
```{r attachment_noRint}
if(!exists("models.logi.attach.noRint") | refresh){
  models.logi.attach.noRint <- glmer(
    lowAttach~Condition_Q*Condition_GP+Reading + 
      (1+Condition_Q*Condition_GP+Reading | SID) + 
      (1+Condition_Q*Condition_GP+Reading | IID),
    data = mdata,
    family = binomial, 
    control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e8)),
    verbose = 2
  )
}
```
```{r attachment_noR}
if(!exists("models.logi.attach.noR") | refresh){
  models.logi.attach.noR <- glmer(
    lowAttach~Condition_Q*Condition_GP + 
      (1+Condition_Q*Condition_GP | SID) + 
      (1+Condition_Q*Condition_GP | IID),
    data = mdata,
    family = binomial, 
    control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e8)),
    verbose = 2
  )
}
```
```{r attachment_noR_r2}
if(!exists("models.logi.attach.noR.r2") | refresh){
  models.logi.attach.noR.r2 <- glmer(
    lowAttach~Condition_Q*Condition_GP + 
      (1+Condition_Q*Condition_GP | SID) + 
      (1+Condition_Q*Condition_GP | IID),
    data = subset(mdata, Reading == 2),
    family = binomial, 
    control = glmerControl(optimizer = "bobyqa", optCtrl=list(maxfun=2e8)),
    verbose = 2
  )
}
```


## Predicting PP1 as strongest or only break

Is PP1 being the strongest or only break predicted by condition * reading?

# Full model

```{r summarizeAttach}
summary(models.logi.attach.full.wD)
```

# Model with no interaction of reading
Simplified model and comparison to full model

```{r sumNoRInt}
summary(models.logi.attach.noRint)
anova(models.logi.attach.full.wD,models.logi.attach.noRint)
```

# Model with no fixed effect of reading

Simplified model and comparison to full model

```{r sumNoR}
summary(models.logi.attach.noR)
anova(models.logi.attach.full.wD,models.logi.attach.noR)
```

# Model with just reading 2

Simplified model and comparison to full model

```{r sumR2}
summary(models.logi.attach.noR.r2)
#anova(models.logi.attach.full.wD,models.logi.attach.noR.r2)
```