---
title: "Prosody description"
author: "Tyler Peckenpaugh"
date: "4/15/2019"
output: 
  bookdown::tufte_handout2: 
    toc: True
    latex_engine: xelatex
---

```{r setup2, include=FALSE}
library(readr)
library(knitr)
library(pander)
library(lme4)
library(stargazer)
library(lattice)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.height = 3)
mdata <- read_csv("~/dissertation/csvs/got_prosody.csv")
mdata$V<-mdata$V == "YES"
mdata$OBJ<-mdata$OBJ == "YES"
mdata$PP1<-mdata$PP1 == "YES"
mdata$STRUG<-mdata$STRUG == "YES"
mdata$condition <- with(mdata,
  paste(
    ifelse(Condition_Q,"+Q","-Q"),
    ifelse(Condition_GP,"+GP","-GP")
  )
)
mdata$STRONG <- as.factor(mdata$STRONG)
mdata$WEAK <- as.factor(mdata$WEAK)
mdata$Reading <- as.factor(mdata$Reading)
mdata$prosody <- as.factor(mdata$prosody)
mdata$two_level_prosody <- as.factor(mdata$two_level_prosody)
mdata$IID <- as.factor(mdata$IID)
mdata$UID <- as.factor(mdata$UID)
mdata$SID <- as.factor(mdata$SID)
mdata$STRUG_START <- as.factor(mdata$STRUG_START)
mdata$pdom <- mdata$two_level_prosody %in% c("PP1", "PP1 > OBJ")
mdata$odom <- mdata$two_level_prosody %in% c("OBJ", "OBJ > PP1")
gpInsensitive <- paste(c(1,
12,
14,
15,
19,
2,
20,
203,
21,
210,
22,
4,
6))

```

# Breaks by condition

## PP1 Break

```{r pp1breaks, fig.cap="PP1 Break"}
br_rows <- c("No break", "Break")
pp1br <- xtabs(~PP1 + condition, mdata)
row.names(pp1br) <- br_rows
pp1br.r1 <- xtabs(~PP1 + condition, subset(mdata, Reading==1))
row.names(pp1br.r1) <- br_rows        
pp1br.r2 <- xtabs(~PP1 + condition, subset(mdata, Reading==2))
row.names(pp1br.r2) <- br_rows

kable(pp1br, caption="Both readings")
kable(pp1br.r1, caption="Reading 1")
kable(pp1br.r2, caption="Reading 2")
```

## OBJ Break

```{r objbreaks, fig.cap="OBJ Break"}
br_rows <- c("No break", "Break")
pp1br <- xtabs(~OBJ + condition, mdata)
row.names(pp1br) <- br_rows
pp1br.r1 <- xtabs(~OBJ + condition, subset(mdata, Reading==1))
row.names(pp1br.r1) <- br_rows        
pp1br.r2 <- xtabs(~OBJ + condition, subset(mdata, Reading==2))
row.names(pp1br.r2) <- br_rows

kable(pp1br, caption="Both readings")
kable(pp1br.r1, caption="Reading 1")
kable(pp1br.r2, caption="Reading 2")
```

# Breaks by &plusmn;Q

## PP1 Break by &plusmn;Q

```{r gpp1, fig.cap="OBJ Break"}
br_rows <- c("No break", "Break")
pp1br <- xtabs(~PP1 + Condition_Q, mdata)
row.names(pp1br) <- br_rows
pp1br.r1 <- xtabs(~PP1 + Condition_Q, subset(mdata, Reading==1))
row.names(pp1br.r1) <- br_rows        
pp1br.r2 <- xtabs(~PP1 + Condition_Q, subset(mdata, Reading==2))
row.names(pp1br.r2) <- br_rows

kable(pp1br, caption="Both readings",col.names=c("-Q","+Q"))
kable(pp1br.r1, caption="Reading 1",col.names=c("-Q","+Q"))
kable(pp1br.r2, caption="Reading 2",col.names=c("-Q","+Q"))
```

## OBJ Break by &plusmn;Q

```{r gpobj, fig.cap="OBJ Break"}
br_rows <- c("No break", "Break")
pp1br <- xtabs(~OBJ + Condition_Q, mdata)
row.names(pp1br) <- br_rows
pp1br.r1 <- xtabs(~OBJ + Condition_Q, subset(mdata, Reading==1))
row.names(pp1br.r1) <- br_rows        
pp1br.r2 <- xtabs(~OBJ + Condition_Q, subset(mdata, Reading==2))
row.names(pp1br.r2) <- br_rows

kable(pp1br, caption="Both readings",col.names=c("-Q","+Q"))
kable(pp1br.r1, caption="Reading 1",col.names=c("-Q","+Q"))
kable(pp1br.r2, caption="Reading 2",col.names=c("-Q","+Q"))
```

# Breaks by &plusmn;GP

## PP1 Break by &plusmn;GP

```{r qpp1, fig.cap="OBJ Break"}
br_rows <- c("No break", "Break")
pp1br <- xtabs(~PP1 + Condition_GP, mdata)
row.names(pp1br) <- br_rows
pp1br.r1 <- xtabs(~PP1 + Condition_GP, subset(mdata, Reading==1))
row.names(pp1br.r1) <- br_rows        
pp1br.r2 <- xtabs(~PP1 + Condition_GP, subset(mdata, Reading==2))
row.names(pp1br.r2) <- br_rows

kable(pp1br, caption="Both readings",col.names=c("-GP","+GP"))
kable(pp1br.r1, caption="Reading 1",col.names=c("-GP","+GP"))
kable(pp1br.r2, caption="Reading 2",col.names=c("-GP","+GP"))
```

## OBJ Break by &plusmn;GP

```{r qobj, fig.cap="OBJ Break"}
br_rows <- c("No break", "Break")
pp1br <- xtabs(~OBJ + Condition_GP, mdata)
row.names(pp1br) <- br_rows
pp1br.r1 <- xtabs(~OBJ + Condition_GP, subset(mdata, Reading==1))
row.names(pp1br.r1) <- br_rows        
pp1br.r2 <- xtabs(~OBJ + Condition_GP, subset(mdata, Reading==2))
row.names(pp1br.r2) <- br_rows

kable(pp1br, caption="Both readings",col.names=c("-GP","+GP"))
kable(pp1br.r1, caption="Reading 1",col.names=c("-GP","+GP"))
kable(pp1br.r2, caption="Reading 2",col.names=c("-GP","+GP"))
```

# 3-level prosodic pattern

```{r 3lvlxtabs}
prosodies.full <- xtabs(~ prosody +condition, mdata)
prosodies.r1 <- xtabs(~ prosody +condition, subset(mdata,Reading==1))
prosodies.r2 <- xtabs(~ prosody +condition, subset(mdata,Reading==2))
kable(prosodies.full, caption = "Both readings")
kable(prosodies.r1, caption = "Reading 1")
kable(prosodies.r2, caption = "Reading 2")
```

# 2-level prosodic pattern

```{r 2lvlxtabs}
prosodies.full <- xtabs(~ two_level_prosody +condition, mdata)
prosodies.r1 <- xtabs(~ two_level_prosody +condition, subset(mdata,Reading==1))
prosodies.r2 <- xtabs(~ two_level_prosody +condition, subset(mdata,Reading==2))
kable(prosodies.full, caption = "Both readings")
kable(prosodies.r1, caption = "Reading 1")
kable(prosodies.r2, caption = "Reading 2")
```

# PP1 or PP1 > OBJ

```{r pdom}
pdom <- xtabs(~condition+pdom,data=mdata)
pdom.r1 <- xtabs(~condition+pdom,data=subset(mdata,Reading==1))
pdom.r2 <- xtabs(~condition+pdom,data=subset(mdata,Reading==2))


kable(
  pdom, caption="Both readings",
  col.names = c("PP1 Not Dominant", "PP1 Dominant")
)
kable(
  pdom.r1, caption="Reading 1",
  col.names = c("PP1 Not Dominant", "PP1 Dominant")
)
kable(
  pdom.r2, caption="Reading 2",
  col.names = c("PP1 Not Dominant", "PP1 Dominant")
)
```

# OBJ or OBJ > PP1

```{r odom}
odom <- xtabs(~condition + odom ,data=mdata)
odom.r1 <- xtabs(~condition + odom,data=subset(mdata,Reading==1))
odom.r2 <- xtabs(~condition + odom,data=subset(mdata,Reading==2))

kable(odom, caption="Both readings",
      col.names = c("OBJ Not Dominant", "OBJ Dominant"))
kable(odom.r1, caption="Reading 1",
      col.names = c("OBJ Not Dominant", "OBJ Dominant"))
kable(odom.r2, caption="Reading 2",
      col.names = c("OBJ Not Dominant", "OBJ Dominant"))
```

# Logistic regression models

The interaction between &plusmn;GP and &plusmn;Q approaches significance (p < 0.06) as a predictor of the object break, but not the PP1 break.

```{r models,results="asis"}
obj <- glmer(OBJ~Condition_GP*Condition_Q+Reading+(1|IID)+(1|SID),data=mdata,family=binomial)
pp <- glmer(PP1~Condition_GP*Condition_Q+Reading+(1|IID)+(1|SID),data=mdata,family=binomial)

stargazer(obj,pp,header=F)
```
