---
title: Methodology
bibliography: refs.bib
output: 
  bookdown::html_document2: 
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.height = 3)
```

This section outlines the methodology employed for the reported study.

# Materials

A number of considerations went into creating the materials for this study. The experimental stimuli are based on earlier pilot study exploring this same phenomenon [@qp2], but with several issues with that study's experimental items corrected.

One of these issues was that the verbs did not all strongly subcategorize for a PP-argument. This could affect the strength of the garden path effect, since the strength of that subcategorization might impact the amount of sway that the parsing strategy of *Minimal Attachment* or *Argument Preference* has on the parse. Theoretically, verbs with stronger subcategorization for a PP-argument would result in a stronger garden path effect, or more difficulty in understanding the sentence. 

Another issue had to do with the level of detail present in the sentences, especially when they were presented as yes/no-questions. When it’s clear that the speaker already knows so much about the situation, it becomes difficult to imagine a pragmatically plausible context where the question would be asked. Such sentences might be described as “prosecutorial[^dcb-prosec].” 

There was also an issue of whether the temporary ambiguity in some of the non-garden path items was ever fully disambiguated. This section will go over considerations taken in order to adjust the stimuli from the earlier study into more systematic and consistent format. The construction of these items is discussed in detail in the rest of this section.

[^dcb-prosec]: Thank you to Dianne Bradley for making this observation, and for the very clever “prosecutorial” descriptor.

## Experimental item versions

The experimental items for this study consisted of 16 sentences in 4 versions: declarative and yes/no interrogative versions of both a PP-attachment garden path of the type discussion earlier, and a non-garden path control which is identical to the garden path up until the final PP, when that PP disambiguates the sentence such that the parse is plausible with a PP2 that modifies the head noun of PP1. The interrogative is always identical to the declarative in every way except for inversion of the order of the subject and auxiliary verb, and the replacement of a final period with a final question mark. An example item in all four versions is shown in \@ref(tab:sentences).

```{r sentences}
library(huxtable)
library(dplyr)
sentences <- list(
  "Declarative non-garden path" = "He had intended to cram the paperwork in the drawer of his filing cabinet.",
  "Interrogative non-garden path" = "Had he intended to cram the paperwork in the drawer of his filing cabinet?",
  "Declarative garden path" = "He had intended to cram the paperwork in the drawer into his boss's desk.",
  "Interrogative garden path" = "Had he intended to cram the paperwork in the drawer into his boss's desk?"
)
tab <- t(hux(sentences,add_colnames = T)) 
names(tab) <- c("Version","Sentence")
tab %>% add_colnames() %>%
  set_top_border(value=1,row=1,col=1:2) %>%
  set_bottom_border(value=1,row=1,col=1:2) %>%
  set_bottom_border(value=1,row=final(1),col=1:2) %>%
  set_caption("Experimental item exemplars") %>%
  set_label("tab:sentences") %>%
  set_italic(TRUE,row=2:5,col=1) 
```

## Main verbs

An earlier pilot study exploring this same phenomenon (Peckenpaugh, 2016) had several issues with its materials. One of these issues was that the verbs did not all strongly subcategorize for a PP-argument. This could affect the strength of the garden path effect, since the strength of that subcategorization might impact the amount of sway that the parsing strategy of Minimal Attachment or Argument Preference has on the parse. Theoretically, verbs with stronger subcategorization for a PP-argument would result in a strong garden path effect. Consider:

  (@weak) Weak PP-argument subcategorization (*hide*) <br>
          a. The gangsters had hidden the shotguns in a U-Haul truck in an old outhouse. <br>
          b. &check; The gangsters had hidden the shotguns.
            
  (@strong) Strong PP-argument subcategorization (*cram*) <br>
            a. Justin had crammed the old newspapers under the couch in a small wastebasket. <br>
            b. * Justin had crammed the old newspapers.
            
The examples in (@weak) and (@strong) point out that while you really must cram something somewhere, you don’t necessarily need to say where you hide something. That is, the verb *hide* seen in (@weak) does not strongly subcategorize for a locative argument (hence the acceptability of (@weak~\tiny ~b)), while cram (@strong) does, rendering (@strong~\tiny ~b) unacceptable. Because of the asymmetry in subcategorization strength, sentences like (@weak) might not trigger as strong a garden-path effect, because the parser might assume that no PP-argument is forthcoming, while for (@strong) the parser will jump at the opportunity to parse a PP as an argument.  In designing the items for this study, only verbs with strong PP-argument subcategorization will be used.

## Fillers

There were 32 filler items that ranged in complexity. Half (16) were designed to end in a sequence of two PPs, to mirror the experimental item. The other half contained no final PPs. All fillers were designed in two versions: declarative and interrogative. A full list of fillers is available in \@ref(appendix-fillers).

## Length

Length was tightly controlled across items. For experimental items, all sentences were between 66 and 75 characters long, and between 13 and 15 words long. The length never varied across the &plusmn;Q condition. Across the &plusmn;GP condition there was a maximum length difference of one character. Two items varied in word length across &plusmn; by one word. For all length variations (word and character), an equal number were longer in the +GP condition as in the -GP condition. They ranged from 18 to 22 syllables.

Control over filler length was slightly less stringent. They ranged from 63 to 79 characters and 12 to 14 words. 

`r if (knitr::is_html_output()) '# References {-}'`

