# Results

This section reports various analyses of the recordings obtained. The analyses include the effect of sentence type (declarative vs. interrogative, garden path vs. non-garden path) on the location of prosodic breaks, as well as "study" or inter-reading time (IRT). In order to evaluate the extent to which participants followed the instruction to read immediately for reading 1 and then study the sentence for reading 2, the amount of time that a sentence is displayed before a participant begins to read it (delay) is also discussed and compared across reading 1 vs. reading 2.

## Participants in the reported analyses

There is a total of 32 participants included, 48 total items, 16 of which are experimental items, with each item read twice by each participant. The 32 filler items are not included in analyses except where specified. The same recordings are used for all analyses.

```{r results_setup,echo=F,warning=F,message=F}
knitr::opts_chunk$set(
  echo=FALSE, 
  warning=FALSE, 
  message=FALSE, 
  fig.height = 3, 
  fig.pos="H"
)
options(tinytex.verbose = TRUE)
bgcolor <- "#ffffff"
par(bg = bgcolor)

kable <- function(data,...) {
  knitr::kable(data, booktabs = TRUE,...)
}

library(readr)
library(huxtable)
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(ggthemes)
library(jsonlite)
library(kableExtra)
library(psych)
library(lme4)

### global plot themeing
theme_set(
  theme_tufte() + 
    theme(
      plot.background = element_rect(fill = bgcolor,color=bgcolor),
      panel.border = element_blank()
    )
)

```

```{r prosody_setup}
mdata_raw <- read_csv("../csvs/got_prosody.csv")
### remove items where STRONG == NA (these are the items Sita didn't respond to)
mdata <- mdata_raw[!is.na(mdata_raw$STRONG),]

### code for pairs of SUBJ-ITEM
mdata$pair<-paste(mdata$SID,mdata$IID,sep="-")
n1 <- nrow(mdata)
### filter out rows missing its pair
mdata<-mdata %>% group_by(pair) %>% filter(n()>1)
n2<- nrow(mdata)
missingPairs <- n1-n2

mdata$V<-mdata$V == "YES"
mdata$OBJ<-mdata$OBJ == "YES"
mdata$PP1<-mdata$PP1 == "YES"
mdata$STRUG<-mdata$STRUG == "YES"
mdata$condition <- with(mdata,
  paste(
    ifelse(Condition_Q,"Q","D"),
    ifelse(Condition_GP,"+GP","-GP")
  )
)
mdata$STRONG <- as.factor(mdata$STRONG)
mdata$WEAK <- as.factor(mdata$WEAK)
mdata$Reading <- as.factor(mdata$Reading)
mdata$prosody <- as.factor(mdata$prosody)
mdata$two_level_prosody <- as.factor(mdata$two_level_prosody)
mdata$IID <- as.factor(mdata$IID)
mdata$UID <- as.factor(mdata$UID)
mdata$SID <- as.factor(mdata$SID)
mdata$STRUG_START <- as.factor(mdata$STRUG_START)
mdata$pdom <- mdata$two_level_prosody %in% c("PP1", "PP1 > OBJ")
mdata$odom <- mdata$two_level_prosody %in% c("OBJ", "OBJ > PP1")
mdata$simple2lvl <- ifelse(
  mdata$OBJ & mdata$PP1, 
  "BOTH", ifelse(
    mdata$OBJ,
    "OBJ",
    ifelse(
      mdata$PP1,
      "PP1",
      "NEITHER"
    )
  )
)
mdata$simple2lvl <- as.factor(mdata$simple2lvl)
mdata<-subset(mdata,SID!=8)
mdata <- droplevels(mdata)
mdata <- subset(mdata,!is.na(OBJ)|!is.na(PP1))

```

```{r irt_setup, cache=T}
itemsDf <- read_csv(
  "../csvs/src/items.csv", 
  col_types = cols(
    OBJ = col_skip(), 
    `P>NP1` = col_skip(), 
    `P>NP2+GP` = col_skip(), 
    `P>NP2-GP` = col_skip(), 
    target = col_character()
  )
)
itemsDf$Item<-itemsDf$target
itemsDf$target <- NULL
### load files
use_old <- FALSE
old<-"../csvs/irt-rvad-4_11.csv"
new<-"../csvs/irt-merged.csv"
raw_file <- read_csv(ifelse(use_old,old,new), 
  col_types = cols(
    Item = col_factor(levels = seq(1,48)), 
    Participant = col_character()
  )
)

irt_data <- raw_file

### P 8 is messed up
irt_data<-subset(irt_data,Participant != 8)

### demographics
pmeta <- read_csv("../csvs/p-meta.csv")
pmeta$Participant <- pmeta$ID
irt_data<-merge(pmeta,irt_data,by="Participant")
irt_data$wkd <- ifelse(irt_data$DOTW == "MONDAY", 1,
                ifelse(irt_data$DOTW == "TUESDAY", 2, 
                ifelse(irt_data$DOTW == "WEDNESDAY", 3, 
                ifelse(irt_data$DOTW == "THURSDAY", 4, 
                5))))



  
irt_data$cond <- paste(
  ifelse(irt_data$Condition_Q,"Q","D"),
  ifelse(
    irt_data$isFiller,
    ifelse(irt_data$Condition_GP,"+PP","-PP"),
    ifelse(irt_data$Condition_GP,"+GP","-GP")
  )
)

irt_data$Q <- ifelse(
  irt_data$Condition_Q,
  "Q",
  "D"
)
irt_data$GP <- ifelse(
  irt_data$Condition_GP,
  "+GP",
  "-GP"
)

irt_data_all <- irt_data ### data before trimming

### exclude 250 < IRTs > 25000
irt_min <- 250
irt_max <- 25000

### note how many IRTs are excluded for implausibility
irt_lo <- subset(irt_data, irt < irt_min)
irt_hi <- subset(irt_data, irt > irt_max)

### subset to trimmed items
irt_data <- subset(
  irt_data, 
  irt > irt_min & irt < irt_max
)

### with fillers
irt_data_allitems <- irt_data
irt_fillers <- subset(irt_data,isFiller)

### experimental only
irt_data <- subset(irt_data,!isFiller)
irt_data <- (merge(itemsDf, irt_data, by = 'Item'))

### make Participant and Item factors with specified ref levels 
irt_data$Participant<-relevel(as.factor(irt_data$Participant),ref=5)
irt_data$Item<-relevel(as.factor(irt_data$Item),ref=1)

### winsorize top/bottom 2.5% of data (for normal distrbution, this is +/- 2sd)
### see: https://sciencing.com/relationship-between-standard-deviations-percentiles-8768703.html
win_trim=0.025
irt_data <- irt_data %>% group_by(Participant) %>% mutate(win_irt = winsor(irt, trim=win_trim))
irt_data$wirt<-irt_data$win_irt

### log 10 transform
irt_data$wirt.log10 <- log10(irt_data$win_irt)


### meta 
json <- read_file("../meta-4_11.json")
meta<-fromJSON(json)
handset <- read.csv("../csvs/_SITA_SET_VALUES.csv")
meta[meta$file %in% handset$Filename,]$agg <- NA
meta[meta$file %in% handset$Filename,]$hpf <- NA
hscount <- sum(handset$Filename %in% meta$file)


write_csv(irt_data_allitems,"export/all_data.csv")
```
```{r normalize sets}
irt_data$ID <- paste0(irt_data$Participant,"-",irt_data$Item)
mdata$ID <- paste0(mdata$SID,"-",mdata$IID)

irt_data <- subset(irt_data, ID %in% mdata$ID)
mdata <- subset(mdata,ID %in% irt_data$ID)
```
```{r describe}
### counts
irt_data$Participant <- as.factor(irt_data$Participant)
irt_data$Participant <- relevel(irt_data$Participant,ref=9)
recs<-nrow(irt_data)
subjs<-length(levels(irt_data$Participant))
items<-length(levels(irt_data$Item))
expectN <- subjs * items


### utterance lengths
if(!use_old){
  ### only available in newer files
  uls <- append(irt_data$`R1 UL`,irt_data$`R2 UL`)
  uls_desc <- describe(uls)
  uls_r1_desc <- describe(irt_data$`R1 UL`)
  uls_r2_desc <- describe(irt_data$`R2 UL`)
}

### note properties
irt_props_all <- describe(irt_data$irt)
irt_props <- describe(irt_data$irt)
wirt_props <- describe(irt_data$win_irt)

### sd by condition table
sdByConditionLong <- aggregate(
  irt_data$win_irt,
  by=list(
    "Q"=irt_data$Q,
    "GP"=irt_data$GP
  ),
  FUN=sd
)

### means by condition table
meansByConditionLong <- aggregate(
  irt_data$win_irt,
  by=list(
    "Q"=irt_data$Q,
    "GP"=irt_data$GP
  ),
  FUN=mean
)
names(meansByConditionLong)[3] <- "mean"
### concise means by condition table
meansByCondition <- meansByConditionLong %>% spread(Q,mean) 
names(meansByCondition)[1] <- "Condition"
### add sd and se to meansByConditionLong
meansByConditionLong$sd<-sdByConditionLong$x
meansByConditionLong$se<-sdByConditionLong$x/sqrt(irt_props$n)


### difference in condition means
diffs<-list(Condition="Increase", `D`=diff(meansByCondition$`D`), `Q`= diff(meansByCondition$`Q`))
meansByCondition<-rbind(meansByCondition, diffs)
cdiff <- meansByCondition$`D`[3]-meansByCondition$`Q`[3]

### participant means by condition
pmeansByCondition <- aggregate(
  irt_data$irt,
  by=list(
    "Condition"=irt_data$cond,
    "Participant"=irt_data$Participant
  ),
  FUN=mean
) %>% spread(Condition,x) 
pmeansByCondition$pattern <- (
  pmeansByCondition$`Q +GP` - pmeansByCondition$`Q -GP` <
  pmeansByCondition$`D +GP` - pmeansByCondition$`D -GP`
) 

### simple P means
spm <- aggregate(
  irt_data$irt,
  by=list(
    "Participant"=irt_data$Participant
  ),
  FUN=mean
) 

 
prosDescrip <- describe(mdata)
tot<-as.data.frame(xtabs( ~ Condition_GP + Condition_Q, data=mdata))

orderMap <- read_csv(
  "../csvs/src/order-p-map.csv",
  col_types = cols(
    ID = col_character(),
    Order = col_character(),
    LIST = col_character()
  )
)

orderMap$Participant <- orderMap$ID
orderMap$ID <- NULL

mdata<-merge(mdata,orderMap,by="Participant")
write_csv(mdata,"export/prosody_data.csv")

versiontab <- mdata %>% 
  filter(!duplicated(Participant)) %>%
  with(table("Version"=LIST,Order)) %>%
  addmargins() 

rownames(versiontab)[1:4]<-paste("Version", rownames(versiontab))
```
```{r vtab}
versiontab %>% kable(
  caption="No. of participants per version-order combination",
  col.names = c("Order 1", "Order 2", "Sum"),
  align="c"
) %>% kable_styling() %>% 
  add_header_above(c(" "=1, "Order"=2, " "=1))
```

There were 4 versions of the experiment, and two possible orderings, with approximately equal number of participants for each version-order combination. Ideally there would be 4 participants per version-order combination, but due to participant attrition and issues during data collection, the distribution is as in Table \ref{tab:vtab}.

Some items are missing due to technical issues during data collection. Other data were excluded due to factors explained later. Ultimately, `r nrow(mdata)` recordings are analyzed (`r nrow(irt_data)` pairs).

## Prosody {#results-prosody}

This subsection presents the results of the independent rater's judgements regarding prosodic boundaries. Data are excluded if either recording from a pair of recordings (reading 1 or reading 2) is missing (there were `r missingPairs` incomplete pairs excluded) because without the other member, it is difficult to determine the extent to which the participant followed instructions (i.e. did not study the sentence before reading 1, did study the sentence before reading 2). Recordings with no reported breaks are excluded from the prosodic analyses, since they are unlikely to represent the reader's intended prosodic structure (there were `r table(mdata$simple2lvl)["NEITHER"]` such recordings). The paired readings of these are not excluded, and IRTs extracted from them are included in the timing analyses, since delay information is available for both readings.

```{r totByCond}
totDisp <- tot %>% as.data.frame() %>%
  spread(Condition_Q,Freq) 
totDisp[[1]] <- c("-GP", "+GP")
kable(
  totDisp,
  col.names=c(" ", "D", "Q"),
  caption="N by condition"
) %>% kable_styling(latex_options = c("hold_position"))
```

Table \ref{tab:totByCond} shows the number of recordings included in analyses for each condition. If every items for each of the 32 participants had been available and included, the expected number per bin would be `r 32*16/4*2`[^n].

[^n]: There are 32 participants, 16 items, and 2 readings, divided across 4 bins.

The sentences can be informally broken into regions as in \ref{@structure}.

  (@structure) [~SUBJ~ She] [~VERBS~ had intended to set] [~OBJ~ the clothes] ~OBJ~ [~PP1~ in the hamper] ~PP1~ [~PP2~ onto the dresser].

In what follows, I will discuss prosodic breaks at the points directly after the direct object (referred to as OBJ) and directly after PP1 (referred to as PP1).

#### Breaks after PP1 and direct object

The PP1 break is expected when PP2, attaches high; i.e., the garden path condition, and is not expected in the non-garden paths, when PP2 attaches low. The OBJ break does not correspond to any expected syntactic structure, and likely represents a breath group or some other not-syntactically-motivated break reason.

```{r byGPall}


pp1tab<-xtabs(PP1 ~ Condition_GP + Condition_Q, data=mdata)
objtab<-xtabs(OBJ ~ Condition_GP + Condition_Q, data=mdata)
tot<-as.data.frame(xtabs( ~ Condition_GP + Condition_Q, data=mdata))


pp1<-as.data.frame(pp1tab)
pp1$tot<-tot$Freq
pp1$pct <- pp1$Freq/pp1$tot
pp1$pretty <- paste0(
  pp1$Freq, 
  " (", 
  sprintf("%.1f",pp1$pct * 100), 
  "%)"
)
pp1disp <- pp1 %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

obj<-as.data.frame(objtab)
obj$tot<-tot$Freq
obj$pct <- obj$Freq/obj$tot
obj$pretty <- paste0(
  obj$Freq, 
  " (", 
  sprintf("%.1f",obj$pct * 100), 
  "%)"
)
objdisp <- obj %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

pp1obj <- cbind(pp1disp,objdisp[2:3])
pp1obj[[1]] <- c("+GP","-GP")
```
```{r pp1gp}

kable(
  pp1obj,
  caption="Percentage of recordings containing PP1/OBJ break by condition",
  col.names = c("", "D", "Q","D", "Q")
) %>% 
  add_header_above(c(" "=1,"PP1 Break"=2,"OBJ Break"=2)) %>% kable_styling(latex_options = c("hold_position"))
```

Table \ref{tab:pp1gp} shows the presence of object and PP1 breaks by condition.

#### The two readings

Because "natural prosody" is likely to be less often produced in reading 1 than in a reading 2, it's important to consider the prosodic patterns for the two readings independently. 

```{r byGPr2}
### r2
r2data <- subset(mdata,Reading==2)

pp1tab<-xtabs(PP1 ~ Condition_GP + Condition_Q, data=r2data)
objtab<-xtabs(OBJ ~ Condition_GP + Condition_Q, data=r2data)
tot<-as.data.frame(xtabs( ~ Condition_GP + Condition_Q, data=r2data))


pp1<-as.data.frame(pp1tab)
pp1$tot<-tot$Freq
pp1$pct <- pp1$Freq/pp1$tot
pp1$pretty <- paste0(
  pp1$Freq, 
  " (", 
  sprintf("%.1f",pp1$pct * 100), 
  "%)"
)
pp1disp <- pp1 %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

obj<-as.data.frame(objtab)
obj$tot<-tot$Freq
obj$pct <- obj$Freq/obj$tot
obj$pretty <- paste0(
  obj$Freq, 
  " (", 
  sprintf("%.1f",obj$pct * 100), 
  "%)"
)
objdisp <- obj %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

pp1obj <- cbind(pp1disp,objdisp[2:3])
pp1obj[[1]] <- c("-GP","+GP")

### r1
r1data <- subset(mdata,Reading==1)

pp1tab<-xtabs(PP1 ~ Condition_GP + Condition_Q, data=r1data)
objtab<-xtabs(OBJ ~ Condition_GP + Condition_Q, data=r1data)
tot<-as.data.frame(xtabs( ~ Condition_GP + Condition_Q, data=r1data))


pp1<-as.data.frame(pp1tab)
pp1$tot<-tot$Freq
pp1$pct <- pp1$Freq/pp1$tot
pp1$pretty <- paste0(
  pp1$Freq, 
  " (", 
  sprintf("%.1f",pp1$pct * 100), 
  "%)"
)
pp1disp <- pp1 %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

obj<-as.data.frame(objtab)
obj$tot<-tot$Freq
obj$pct <- obj$Freq/obj$tot
obj$pretty <- paste0(
  obj$Freq, 
  " (", 
  sprintf("%.1f",obj$pct * 100), 
  "%)"
)
objdisp <- obj %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

pp1objr1 <- cbind(pp1disp,objdisp[2:3])
pp1objr1[[1]] <- c("-GP","+GP")
```

The prevalence of object breaks and PP1 breaks for each condition in reading 2 recordings is described in table \ref{tab:r2pp1obj}.

```{r r2pp1obj}
kable(
  pp1obj,
  caption="Reading 2 recordings containing PP1/OBJ break by condition",
  col.names = c("", "D", "Q","D", "Q")
) %>% 
  add_header_above(c(" "=1,"PP1 Break"=2,"OBJ Break"=2)) %>% 
  kable_styling(latex_options = c("hold_position"))
```

For comparison, \ref{tab:r1pp1obj} is the same table for reading 1 recordings. 

```{r r1pp1obj}
kable(
  pp1objr1,
  caption="Reading 1 recordings containing PP1/OBJ break by condition",
  col.names = c("", "D", "Q","D", "Q")
) %>% 
  add_header_above(c(" "=1,"PP1 Break"=2,"OBJ Break"=2)) %>% 
  kable_styling(latex_options = c("hold_position"))
```

The primary difference across readings is in the OBJ break distribution. For the analyses that follow, reading 1 will largely be ignored.

#### PP1 and object breaks combined

It is perhaps more interesting to look at the combined distribution of the breaks: i.e., a sentence can have one of four patterns: OBJ and PP1 breaks are present; only OBJ or only PP1 is present; or neither is present. As mentioned earlier, there were only 5 cases where neither are present, and those are excluded.

```{r r2combobreaksBycondpre}
r2pdata<-subset(mdata,Reading==2)
qdata <- subset(r2pdata,Condition_GP)
ddata <- subset(r2pdata,!Condition_GP)
qpros<-table(qdata$simple2lvl,qdata$Condition_Q) %>%
  prop.table(margin=2) %>% `*`(100) %>%
  round(1)
colnames(qpros)<-c("D","Q")

dpros<-table(ddata$simple2lvl,ddata$Condition_Q) %>%
  prop.table(margin=2) %>% `*`(100) %>%
  round(1)
colnames(dpros)<-c("D","Q")

ccombo<-cbind(dpros,qpros)
```
```{r r2combobreaksBycond}
kable(ccombo,caption="Combined breaks by condition in reading 2 only") %>%
  add_header_above(c(" "=1,"Non-garden path" = 2, "Garden path"=2)) %>%
  column_spec(1,bold=T)%>% kable_styling(latex_options = "hold_position")

```

Table \ref{tab:r2combobreaksBycond} shows that for +GP sentences, there are very few instances with the OBJ-only pattern (0.8% in declaratives, 2.5% in interrogatives); whereas that pattern is fairly robust for -GP sentences (31.1% in declaratives, 31.4% in interrogatives). The pattern with both breaks is somewhat more common for +GP sentences (72.1% in declaratives, 71.7% in interrogatives) than -GP (54.1% in declaratives, 43.0% in interrogatives). The PP1-only pattern occurs at about the same rate in +GP interrogatives (25.8%) as in -GP interrogatives (25.6%) and +GP declaratives (27%), but is noticeably less common for -GP declaratives (14.8%).

#### PP1 and object breaks and their relative prominence

Table \ref{tab:pros2lvl} presents data similar to those in the previous section, but it incorporates the rater's judgement of the relative prominence of the breaks. The &gt; symbol indicates that the rater found the break on the left of that symbol to be stronger, or more prominent, than the break on the left. When no symbol is shown between the two breaks, the rather found them to be of equal prominence. Please be aware that inter-rater reliability for relative prominence was not good, as discussed in section \ref{raterRel}.

```{r pros2lvlPre}

### TODO present this table across condition(s)

allpros<-table(mdata$two_level_prosody)%>%
  prop.table() %>% 
  `*` (100) %>%
  round(1) %>%
  hux() %>%
  setNames(c("Pattern", "Percent of all data"))
r2pros<-table(
  subset(mdata,Reading==2)$two_level_prosody
) %>%
  prop.table() %>% 
  `*` (100) %>%
  round(1) %>%
  hux() %>%
  setNames(c("Pattern", "r2")) 
r1pros<-table(
  subset(mdata,Reading==1)$two_level_prosody
) %>%
  prop.table() %>% 
  `*` (100) %>%
  round(1) %>%
  hux() %>%
  setNames(c("Pattern", "r1")) 

allpros$`Cold readings` <- r1pros$r1
allpros$`Previewed readings` <- r2pros$r2
```
```{r pros2lvl}
kable(allpros,caption="Prosodic pattern by reading")%>% kable_styling(latex_options = "hold_position")
```

\clearpage

### Regression models predicting break occurence

Models predicting the PP1 break and the object break are presented in table \ref{tab:lmPros}. No random slopes were included, as those models did not converge.

```{r prosmodelsPre,cache=T}
pp1Mod<- glmer(PP1~Reading+Condition_Q*Condition_GP+(1|IID)+(1|SID),data=mdata,family=binomial)
objMod<- glmer(OBJ~Reading+Condition_Q*Condition_GP+(1|IID)+(1|SID),data=mdata,family=binomial)
```
```{r prosmodels}
huxreg(  
  list("PP1"=pp1Mod,"Object"=objMod),
  coefs = c(
    "Intercept"="(Intercept)",
    "GP"="Condition_GPTRUE",
    "Q"="Condition_QTRUE",
    "GP:Q Interaction"="Condition_QTRUE:Condition_GPTRUE",
    "Preview" = "Reading2"
  ),
  statistics = c(
    N="nobs",
    "logLik",
    "AIC"
  )
) %>%
  set_caption("Logistic regression models of prosody") %>%
  set_label("lmPros") %>%
  set_tabular_environment("tabular")
```

\clearpage

## Inter-reading time (IRT) {#results-irt}


This document examines the inter-reading time (IRT) from the study. IRT was measured over `r recs` recordings: 32 participants, 48 items = `r subjs * items` recording pairs (reading 1 and reading 2), with `r  subjs * items - recs` missing pairs. The missing data are a result of one or both recordings from a pair being unusable due to technical issues (e.g. a failure of recording equipment, or participant error).

#### Distribution of IRT

The raw IRTs including fillers and before any outliers are trimmed are distributed as shown in Figure \ref{fig:rawIRThist}. Overall mean IRT of these data (n = `r irt_props_all$n`), is `r round(irt_props_all$mean/1000,1)`s. The longest is `r round(irt_props_all$max/1000,1)`s and the shortest `r round(irt_props_all$min,0)`ms. Median IRT is `r round(irt_props_all$median/1000,1)`s.

```{r rawIRThist,fig.cap="Distribution of raw IRT"}
histSettings = geom_histogram(binwidth = 0.5,color="white",fill="#333333")
ggplot(irt_data_all, aes(irt/1000)) +
  histSettings +
  xlab("Raw IRT (seconds)") + 
  ylab("Frequency") +  
  ggtitle(
    "Distribution of raw IRT", 
    subtitle="Bin size = 0.5s"
  )
```

IRTs below `r irt_min`ms (`r nrow(irt_lo)`) and above `r round(irt_max/1000,1)`s (`r nrow(irt_hi)`) are (assumed to be implausible) omitted. Experimental data were then Winsorized by participant to bring data in the `r win_trim*100`th and `r 100-(100*win_trim)`th percentile of data to the value at those tresholds. The resulting measure is referred to as wIRT and is distribued as shown in Figure \ref{fig:wIRT} (n = `r wirt_props$n`). Overall mean for wIRT is `r round(wirt_props$mean/1000,1)`s. The longest IRT is `r round(wirt_props$max/1000,1)`s and the shortest is `r round(wirt_props$min,0)`ms. Median wIRT is `r round(wirt_props$median/1000,1)`s.

```{r wIRT,fig.cap="Distribution of wIRT"}
ggplot(irt_data, aes(win_irt/1000)) +
  histSettings + 
  xlab("wIRT (seconds)") + 
  ylab("Frequency") + 
  ggtitle(
    "Distribution of wIRT", 
    subtitle="Bin size = 0.5s"
  )
```
\clearpage

#### Means by condition

Table \ref{tab:mns} shows the mean wIRT by experimental condition. The top left cell represents the mean wIRT for the declaritive controls ("D -GP"). The bottom row shows the increase in IRT across the garden path condition. 

```{r mnspre}
meansByConditionS <- meansByCondition
meansByConditionS$`D` <- meansByConditionS$`D`/1000
meansByConditionS$`Q` <- meansByConditionS$`Q`/1000
```
```{r mns}
kable(
  meansByConditionS, 
  caption = "Means (s) by condition",
  digits = 2
)%>% kable_styling(latex_options = "hold_position")
```

The difference in the effect of &plusmn;GP across &plusmn;Q is `r round(cdiff/1000,2)`s. That is, the mean amount that IRT increased for a garden path declarative compared to a non-garden path declarative is `r round(cdiff/1000,2)`s `r ifelse(cdiff > 0, paste("more"),paste("less"))` than the amount that IRT increased for a garden path interrogative compared to a non-garden path interrogative.

```{r interactinplot,fig.cap="Mean IRT by condition",fig.height=4}

meansByConditionLong$Qn <- ifelse(
  meansByConditionLong$Q == "Q", 
  "Interrogative", 
  "Declarative"
)

meansByConditionLong$GPn <- ifelse(
  meansByConditionLong$GP == "+GP", 
  "Garden path", 
  "Non-garden path"
)

ggplot(
  meansByConditionLong, 
  aes(mean/1000,x=GPn,y=mean/1000,group=Qn,linetype=Qn)
) + 
  geom_line() +
  geom_point() +
  ylim(6,7.2) +
  geom_errorbar(aes(ymin=(mean-se)/1000, ymax=(mean+se)/1000), width=.125) +
  labs(
    x="", 
    y="Mean IRT (ms)",
    linetype="",
    title="Mean IRT by condition", 
    subtitle = "Confidence intervals represent one standard error"
  )
```

\clearpage

#### Regression models of IRT

The models with random slopes for participant and item did not converge, so the tables in this section show models with no random slopes. 

```{r lmeMods, cache=T}

irt.full <- lme4::lmer(
  wirt ~ Condition_Q * Condition_GP + 
    (1 | Participant) + 
    (1 | Item),
  data = irt_data,
  REML=F
)

irt.noInteraction <- lme4::lmer(
 wirt ~ Condition_Q + Condition_GP + 
    (1 | Participant) + 
    (1 | Item),
  data = irt_data,
  REML=F
)

irt.noParticipant <- lme4::lmer(
  wirt ~ Condition_Q * Condition_GP + 
    (1 | Item),
  data = irt_data,
  REML=F
)

irt.noItem <- lme4::lmer(
 wirt~ Condition_Q * Condition_GP + 
    (1 | Participant),
  data = irt_data,
  REML=F
)

irt.noFxd <- lme4::lmer(
 wirt ~
    (1 | Participant) + 
    (1 | Item),
  data = irt_data,
  REML=F
)

irt.noRand <- lm(wirt ~ Condition_Q * Condition_GP,
                 data = irt_data)

irt.dummy <- lm(wirt ~ Condition_Q * Condition_GP,
                  data = irt_data
                )
irt.dummy.noInt <- lm(wirt ~  Condition_Q + Condition_GP + 
                  Participant + Item,
                  data = irt_data
                )

irt.demoModel.KS <- lmerTest::lmer(
  wirt ~ Condition_Q * Condition_GP + Semester + wkd + Gender + ease + timeslot +
    bookReading + lightReading + Monolingual + (1|Participant) + (1|Item), 
  data = irt_data,
  REML=F
)
irt.demoModel <- lmerTest::lmer(
  wirt~ Condition_Q * Condition_GP + Semester + wkd +
    Monolingual + (1|Participant) + (1|Item), 
  data = irt_data,
  REML=F
)

irt.summer <- lmerTest::lmer(
  wirt ~ Condition_Q * Condition_GP +
    Monolingual + (1|Participant) + (1|Item), 
  data = subset(irt_data, Semester!="Summer"),
  REML=F
)
```

For the first model, fixed effects of &plusmn;GP and &plusmn;Q as well is the interaction between them were included, along with random effects of participant and item. The second model removes the interaction, but keeps both main effects.

```{r modTabLME}
huxreg(
  list(
    "Full"=irt.full, 
    "No interaction"=irt.noInteraction,
    "No random effects"=irt.noRand,
    "No fixed effects"=irt.noFxd
  ),
  coefs = c(
    "+GP"="Condition_GPTRUE",
    "Q"="Condition_QTRUE",
    "+GP Q"="Condition_QTRUE:Condition_GPTRUE"
  ),
  note=NULL,
  statistics = c(N="nobs","logLik","AIC")
) %>% 
  set_caption("Models") %>% 
  set_label("tab:models") %>%
  set_tabular_environment("tabular")
```

A model with no fixed effects and one with no random effects were also run. The estimates from these models can be seen in table \ref{tab:models}. 

\clearpage

Several model comparisons were made, seen in tables \ref{tab:modCom3}, \ref{tab:modCom}, and \ref{tab:modCom2}.

```{r modCom3}
cols <- c(
  "Df",
  "AIC",
  "BIC",
  "logLik",
  "deviance",
  "$\\chi^2$",
  "Df($\\chi$)",
  "Pr(>$\\chi^2$)"
)
comp <-anova(irt.full,irt.noFxd)
row.names(comp) <- c("No fixed effects", "Full")
kable(comp,caption="Full model vs. no fixed effects",col.names=cols,escape=F)
```

There are clearly main effects of both &plusmn;GP and &plusmn;Q: table \ref{tab:modCom3} shows that the full model is significantly better than the one with no fixed effects. 

```{r modCom}

comp <-anova(irt.full,irt.noInteraction)
row.names(comp) <- c("No interaction", "Full")
kable(comp,caption="Full model vs. no interaction",col.names=cols,escape=F)

```

The interaction between main effects, though, is not able to be confirmed (in fact, table \ref{tab:modCom} shows the non-interaction to be better, but not to a stastically significance degree).

```{r modCom2}
comp <- anova(irt.full,irt.noRand)
row.names(comp) <- c("No random effects", "Full")

kable(comp,caption="Full model vs. no random effects",col.names=cols,escape=F)
### tab <- hux(comp[,c(1:2,4,6:8)],autoformat = T)
### tab <- add_rows(
###   tab, 
###   hux("Df","AIC","logLik","&Chi;<sup>2</sup>","Df(&Chi;)","p value"),
###   after=0
### )
### tab <- insert_column(
###   tab, c("","No random effects", "Full")
### )
### 
### tab %>% set_escape_contents(F,row=1,col=5:7) %>%
###   set_number_format(value=2,col = c(3:5,7),row=2:3) %>%
###   set_label("tab:modComp2") %>% 
###   set_caption("No random effects vs. full model") %>%
###   set_top_border(row=1,value=1,col=everywhere) %>%
###   set_bottom_border(row=1,value=1,col=2:7) %>% 
###   set_bottom_border(row=final(1),value=1,col=everywhere) 
```

Table \ref{tab:modCom2} shows that the random effects of participant and item improve the model in a stastically signifcant way.

\clearpage

#### Effect of verb on IRT

If we consider the mean wIRT by which of the 4 verbs occured in the target sentence (*cram*, *set*, *put*, or *stick*), we see that there was some difference in how the two experimental manipulations (&plusmn;GP = garden path status and &plusmn;Q = interrogative vs. declarative status) effected wIRT. The means (and standard deviation) across the conditions are reported for each verb in table \ref{tab:verbConMeans}.

```{r verbMeans}
### means by condition table
meansByVerb <- aggregate(
  irt_data$win_irt/1000,
  by=list(
    "Condition"=irt_data$cond,
    "Verb"=irt_data$V
  ),
  FUN=mean
) %>% spread(Condition,x)
### sd by condition table
sdByVerb <- aggregate(
  irt_data$win_irt/1000,
  by=list(
    "Condition"=irt_data$cond,
    "Verb"=irt_data$V
  ),
  FUN=sd
) %>% spread(Condition,x)
meansByVerb$`Declarative GP effect`<-meansByVerb$`D +GP`-meansByVerb$`D -GP`
meansByVerb$`Interrogative GP effect`<-meansByVerb$`Q +GP`-meansByVerb$`Q -GP`
meansByVerb$`Difference in effect`<-meansByVerb$`Declarative GP effect`-meansByVerb$`Interrogative GP effect`
```

```{r verbConMeans}
meansNsd<-sdByVerb
for (col in colnames(meansNsd)[2:5]){
  meansNsd[[col]]<- paste(
    round(meansByVerb[[col]],1), 
    " (",
    round(meansNsd[[col]],1),
    ")",
    sep=""
  )
}

kable(
  meansNsd,
  caption="Mean (sd) wIRT by condition and verb in seconds"
) %>% 
  kable_styling(latex_options = "hold_position")
```

To isolate the effect of the garden path for interrogatives as compared to declartives, we can subtract the mean for declarative non-garden paths from the mean for declarative garden paths, and then do the same for interrogatives. These measures are referred to as the "declarative GP effect" and "interrogative GP effect" in \ref{verbConDiffMeans}. The difference between the declarative GP effect from the interrogative GP effect is labeled the "difference in effect." We can see that for 2 of the 4 verbs (*set* and *stick*), there was a positive difference in effect, while for the other two we find a negative one.

```{r verbConDiffMeans}
kable(meansByVerb[,c(1,6:8)],caption="Effect of GP on wIRT by verb", digits=1) %>% kable_styling(latex_options = "hold_position")
```
#### Individual variation in IRT

Individuals vary with regard to the effect of the garden path condition on IRT. For `r table(pmeansByCondition$pattern)["TRUE"]` of `r nrow(pmeansByCondition)`, the increase in IRT for garden paths is greater for interrogatives than it is for declaratives.

```{r pPattern}
kable(
  pmeansByCondition[c(1:5)],
  caption="Mean wIRT (ms) by condition and participant",
  longtable=T
)%>% kable_styling(latex_options = c("hold_position", "repeat_header"))
```

\clearpage

#### Interrogative processing cost is represented in IRT

Interrogatives appear to have a computational processing cost when compared to interrogatives [cf. @lehiste1973phonetic]. The filler sentences in this study were designed so as to provide a diagnostic of the interrogative effect on IRT. 

```{r fillers, results="asis"}
mean.irtByQ <- aggregate(
  irt_fillers$irt, 
  by=list("Q"=irt_fillers$Condition_Q),
  FUN=mean
)
mean.irtByPP <- aggregate(
  irt_fillers$irt, 
  by=list("+PP"=irt_fillers$Condition_GP),
  FUN=mean
)
mean.irtByQnPP <- aggregate(
  irt_fillers$irt, 
  by=list("Q"=irt_fillers$Condition_Q,"+PP"=irt_fillers$Condition_GP),
  FUN=mean
)
fillerModel <- lme4::lmer(
  irt ~ Condition_Q + (1 |Participant) + (1 |Item), 
  data=irt_fillers, 
  REML=F
)
fillerModel.pp <- lme4::lmer(
  irt ~ Condition_Q * Condition_GP + (1|Participant) + (1 |Item), 
  data=irt_fillers,
  REML=F
)
fillerModel.ppNoInt <- lme4::lmer(
  irt ~ Condition_Q + Condition_GP + (1|Participant) + (1 |Item), 
  data=irt_fillers,
  REML=F
)

huxreg(
  list("Only Interogitivity"=fillerModel,"Full"=fillerModel.pp,"No interaction"=fillerModel.ppNoInt),
  number_format = 2, 
  coefs=c(
    "(Intercept)"="(Intercept)",
    "Q" = "Condition_QTRUE",
    "+GP" ="Condition_GPTRUE",
    "Q * +GP" = "Condition_QTRUE:Condition_GPTRUE" 
  )
) 
```

The mean IRT for interrogative fillers was `r round(mean.irtByQ[1,]$x/1000,1)`s; for declarative, `r round(mean.irtByQ[1,]$x/1000,1)`s. Interrogatives elicited a mean IRT of `r round(diff(mean.irtByQ$x),0)`ms longer than declaratives. Half of the fillers had a string of two PPs at the end. This did not impact IRT to the same degree (a difference of `r abs(round(diff(mean.irtByPP$x),0))`ms across &plusmn;PP).

A mixed effects regression model (no random slope structure, due to convergence errors) found a significant main effect of &plusmn;Q (estimate = 379ms, t = 2.24, p < 0.05)

\clearpage

## Delay comparison for cold vs. previewed readings

A comparison of the delay for cold readings compared with that of previewed readings can lend insight into the extent to which subjects followed task instructions.

"Delay" here is the amount of time after the start of a recording until the beginning of phonation of the target sentence. Cold readings are also called "reading 1", while previewed readings are the same as "reading 2". Implausible delays of >15s are excluded in the data shown here.

```{r delayComparison, fig.cap="Delay comparison across reading"}
raw_rs_file <- read_csv("../csvs/merged.csv")
raw_rs_file$UID <- as.factor(with(raw_rs_file, paste(Participant,Item,Reading,sep="-")))
raw_rs_file <- subset(raw_rs_file,UID %in% mdata$UID)
raw_rs <- subset(raw_rs_file,!isFiller)
raw_rs$reading <- raw_rs$Reading
raw_rs$Reading <- ifelse(raw_rs$reading == 1, "Cold", "Previewed")
ggplot(raw_rs, aes(Leading/1000, fill = Reading)) +
  geom_histogram(binwidth = 0.5,position="dodge",color="black") +
  scale_fill_manual(values=c("black","white")) +
  ggtitle("Distribution of delay by reading", 
          subtitle = "Bin size = 0.5s")

diffs <- raw_rs[c("Reading","Leading","Participant","Item","isFiller",
                  "Condition_Q","Condition_GP")]  %>% spread(Reading,Leading)

diffs$diffs <- diffs$Previewed-diffs$Cold
diffDis<- describe(diffs$diffs)
diffsByP<-aggregate(diffs~Participant,data=diffs,FUN=function(x){round(mean(x))})
colnames(diffsByP) <- c("Participant", "Mean difference in delay (ms)")
diffsByP.props <- describe(diffsByP$`Mean difference in delay (ms)`)
```

For each reading, n = `r table(raw_rs$Reading)["Cold"]` (total n = `r nrow(raw_rs)`).

### Reading 1 delays by participant

```{r}
r1data <- subset(raw_rs, reading == 1)
r1delByP <- aggregate(r1data$Leading,by=list("Participant" = r1data$Participant), FUN=mean)
hidel <- mean(r1delByP$x) + sd(r1delByP$x)/2
lodel <- mean(r1delByP$x) - sd(r1delByP$x)/2
r1delByP$r1DelCat <- ifelse(r1delByP$x < lodel, "FAST", ifelse(
  r1delByP$x > hidel, "SLOW", "NORMAL"
))

colnames(r1delByP)[2] <- "r1delbyp"

irt_data<-merge(r1delByP,irt_data,by="Participant")
mdata$Participant <- mdata$SID
mdata<-merge(r1delByP,mdata,by="Participant")
write_csv(irt_data,"export/exp_wins_irt.csv")
```
```{r}
del.irt.full <- lmerTest::lmer(
  wirt ~  Condition_GP + (1 | Participant) + 
    (1 | Item),
  data = irt_data,
  REML=F
)

mmdata <- subset(mdata,r1DelCat != "NORMAL" & Reading == 2)
del.pp1Mod<- glmer(PP1~Condition_Q*Condition_GP*r1DelCat+
                     (1|IID)+(1|SID),data=mmdata,family=binomial)
del.objMod<- glmer(OBJ~Condition_Q*Condition_GP*r1DelCat+
                     (1|IID)+(1|SID),data=mmdata,family=binomial)
summary(del.irt.full)
summary(del.pp1Mod)
summary(del.objMod)
```

### Difference in delay across paired readings

Overall, each recording pair (n = `r diffDis$n`) has a mean difference in delay (DelDif = previewed delay - cold delay) of `r round(diffDis$mean/1000,1)`s (sd = `r round(diffDis$sd/1000,1)`s), with a minumum of `r round(diffDis$min/1000,1)`s and a max of `r round(diffDis$max/1000,1)`s. The median DelDif is `r round(diffDis$median/1000,1)`s. The distribution DelDif is shown in Figure \ref{fig:deddif}.

```{r deddif, fig.cap="Distribution of DelDif"}
ggplot(na.omit(diffs), aes(diffs)) +
  geom_histogram(binwidth = 250,color="white",fill="black") +
  labs(x="Difference in delay", y="Count",
       title="Distribution of difference in delay", subtitle="Bin width = 250ms")
```

If we calculate the mean delay difference by participant, we find a mean participant DelDef of `r round(diffsByP.props$mean/1000,1)`s. Each participant's DelDif is &le; `r diffsByP.props$min`ms and &ge; `r round( diffsByP.props$max/1000,1)`s, with a median of `r round(diffsByP.props$median/1000,2)`s. Table \ref{tab:difsbyp} shows these values.

```{r difsbyp}
kable(
  diffsByP[order(diffsByP[,2]),],
  row.names=F,
  digits = 0,
  caption="Delay differences by participant",
  longtab=T
)%>% kable_styling(latex_options = c("hold_position", "repeat_header"))
```

The distribution of the participants' DelDifs can be found in Figure \ref{fig:difhistbyp}.

```{r difhistbyp, fig.cap="Mean difference in delay by participant"}
ggplot(diffsByP, aes(`Mean difference in delay (ms)`/1000)) +
  geom_histogram(binwidth = 0.5,color="white",fill="black") +
  labs(x="Mean difference in delay", y="Count",title="Mean difference in delay by participant", subtitle="Bin size = 0.5s")
```
\pagebreak
