---
always_allow_html: yes
---
```{r disc,echo=F,warning=F,message=F}
knitr::opts_chunk$set(
  echo=FALSE, 
  warning=FALSE, 
  message=FALSE, 
  fig.height = 3, 
  fig.pos="H"
)
options(tinytex.verbose = TRUE)
bgcolor <- "#ffffff"
par(bg = bgcolor)

kable <- function(data,...) {
  knitr::kable(data, booktabs = TRUE,...)
}

library(readr)
library(huxtable)
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(ggthemes)
library(jsonlite)
library(kableExtra)
library(psych)
library(lme4)

### global plot themeing
theme_set(
  theme_tufte() + 
    theme(
      plot.background = element_rect(fill = bgcolor,color=bgcolor),
      panel.border = element_blank()
    )
)
ccombo <- read_csv("export/ccombo.csv")
rownames(ccombo)<-c("Both", "OBJ", "PP1")
```
# Discussion and conclusion

This section will go through the motivating questions behind this study and discuss the extent to which those questions are answered, or not. It will discuss the issues that may have lead to the disatisfying. It will then go on to present further questions and propose further studies to explore those new questions and the ones left unanswered here.

## The intuition

Recall that the primary motivator for this study was the observation that PP-attachment garden paths appear easier to understand for many speakers of American English when presented in the interrogative, as opposed to the declarative. One goal of this study was to find a behavioral correlate of this intuition. Another was to seek an explanation for the intuition itself. 
  
### Behavioral correlate for the intuition

Ultimately the data here are unable to provide a clear answer to whether or not IRT represents a behavioral correlate of the intuition that PP-attachment garden paths are rendered easier to parse when presented as polar questions. While the difference in mean IRT for Q+GP compared to Q-GP was 0.04s less than for D+GP compared to D-GP, showing a smaller increase in processing time for interrogatives than for declaratives, this is a rather small difference when the grand mean is 6.6s. As such, mixed-effect regression analyses were not able to detect stastical significance for the interaction between interogativity and &plusmn;GP. That said, it was also not clear that a model without the interaction was a better fit than one with it (Full model AIC = 9103.5; Non-interaction model AIC = 9101.6; &Chi;^2^ = 0.035; p > 0.8). This does not, of course, negate the intuition; it simply means that we have not yet found a behavior that can be said with certainty to correspond to that intuition.

There are a number of issues that might be preventing the observation of behavior that confirms the intuition, and there are also possible explanations for why the intuition might exist without a behavior correlate. The possibilty most generous to the reported study is that the sample size here is simply inadequate. It's possible that with more data, this very procedure would have provided a stastically significant behavioral correlate of the intuition. 

It is also possible that the experimental paradigm used here is too lenient to capture the behavior it is looking for: it relies on participants to begin Reading 1 without any lookahead simply by their own recognizance rather than any experimenter-enforced mechanism. Perhaps some participants were able to, perhaps unintentionally, get enough information a priori to negate the effect of the interaction on their IRT.

It could also be that IRT is simply the wrong measure for detecting a behavioral correlate of the intuition. The reanalysis triggered by +GP Reading 1 could very well be all that was needed to make the understanding of +GP Reading 2 trivial; this seems unlikely, though, because a significant effect of +GP on IRT was found.

### Explaining the intuition

A possible explanation for the intuitive effect of interogativity on parsing garden paths is provided in the work by @Bader1998-ts. Bader demonstrates that it is easier to recover a parse that “behave[s] alike prosodically” to a given failed parse, because the reanalysis does not require prosodic reconstruction and only the syntax needs to be repaired. In the case of the intuition this study is concerned with, this would mean that should sentences be more similar across the garden path vs. non-garden path condition in the interrogative than in the declarative, the intuitive reduction in difficulty of reanalysis would naturally follow. 

While the data presented above do not definitively show that the prosodic structure of questions is less different for questions than it is for declaratives, it also does not rule it out. Because there does not appear to be a one-to-one mapping of prosodic structure to syntactic structure in the recordings, one cannot make a clear-cut declaration of what a given sentence type looks like prosodically. Instead, we must observe the percentage of recordings for each version of the sentences that exhibit a given pattern. The relevant table is repeated here as \@ref(tab:combobr) for convenience.

```{r combobr}
kable(ccombo,caption="Combined breaks per condition in reading 2 only", col.names=c("D","Q","D","Q"),align="c") %>%
  add_header_above(c(" "=1,"Non-garden path" = 2, "Garden path"=2)) %>%
  column_spec(1,bold=T) %>% kable_styling(latex_options = "hold_position")
```

While there is a larger drop in the number of utterances with both breaks from +GP to -GP for questions than for decaratives, the opposite is true for the pattern with a PP1 break by itself. There is an argument to be made that the OBJ break, which does not correspond to a change in syntactic branching direction for any version of the sentences, is a hesitation rather than a break, due to a moment of confusion or perhaps simply a need to breathe. 

To be more explicit, for this explanation to work, the prosodic structures in @dlo and @dhi would likely be the ones considered "correct" or most common for a the declarative versions. The symbol "%" is being used to represent a prominent prosodic break.

  (@dlo) *D+GP*: He had crammed [~OBJ~ the newspapers] [~PP1~ under the sofa] % [~PP2~ into the trashcan].
  (@dhi) *D-GP*: He had crammed [~OBJ~ the newspapers] under the sofa in the guestroom.

The prosodic structure of @dhi does not mandate any major break, because there is no change in syntactic branching directory and no other reason to mandate a break there. That is, the PP1 break differentiates the two syntactic structures and signals the attachment site of PP2. 
 
For the interrogatives, though, this contrast is obscured by the need to apply a rising prosodic contour over the final nuclear accent in the sentence: i.e. PP2, resulting in not a change in syntactic branching direction, but a tonal change, which might sound very similar to a prosodic break. The symbol "//" indicates the start of the rising contour.

  (@qlo) *Q+GP*: Had he crammed [~OBJ~ the newspapers] [~PP1~ under the sofa] %+// [~PP2~ into the trashcan]?
  (@qhi) *Q-GP*: Had he crammed [~OBJ~ the newspapers] [~PP1~ under the sofa]    // [~PP2~ in the guestroom]?

The critical issue here is that in (@qhi), the absence of a syntactically-motivated break is hidden by the juncture created by the start of the final-rising contour of a question. 

If the PP1 break is treated as the main indicator of the prosodic structure, then the larger difference across &plusmn;GP for declaratives (`r 27-14.8`%) than for interrogatives (`r 25.8-25.6`%) does indeed, though perhaps weakly, leave open the door for the @Bader1998-ts style explanation of the intuition.

## On future work {#confound}

In this section, I will discuss some of the issues with the current study, and what I see as fruitful avenues for future study.

### Hesitations vs. prosodic breaks

A major roadblock to the success of the methodology here reported is that the region of the sentence where the critical prosodic difference between +GP and -GP sentences lies is the same region where disambiguation occurs for the +GP sentences. Some weaknesses in the data as they are leave need for future studies and  methodological improvements. 

A major issue is the difficulty in differentiating between syntactically-motivated prosodic breaks and hesitations. As table \@ref(tab:combobr) and the reliability data reported earlier in this paper illustrates, a linguistically trained judge will not necessarily be able to distinguish between prosodic breaks and hesitations, and will not necessarily agree with another linguistically trained judge on the relative prominence of breaks and hesitations within a recording. It is famously difficult to instrumentally describe a prosodic break, and there is no guarentee that examining the wave forms would fare any better in making this distinction than the intuition of the judges. That said, it may be possible to distinguish the two sorts of breaks: it may be that boundary tones or segmental lengthening are typically not part of a hesitation break, but are a part of hesitation break. No immediately apparent pattern is available, but it would be inordinately beneficial if one could be found.

In \@ref(erp) I propose that an event-related potential paradigm might hold the key for distinguishing between the two.

### Embedded questions

The explanation for the intuition that PP-attachment garden paths are less difficult to parse in interrogatives than in declaratives must either be prosodic, or else semantic/pragmatic. It seems impossible that the  very minor syntactic difference across the two sentence types (subject-auxiliary inversion) could be the reason for a difference in percieved ease of understanding. It is either the difference in the intonational contour between the two sentences types, or else it is something about the meaning or metalinguistic difference between interrogatives and declaratives. The study just reported has not provided adequately convincing evidence that the inuition can or can not be explained entirely by the prosodic differences between questions and declaratives. It seems that the next step in explaining the aforementioned intuition would be to look at the same phenomenon in *embedded* questions vs. embedded declarative clauses, where prosody would not be at play. For example:

  (@embed_qgp) *EmQ +GP:* He asked her if she had decided to cram the old newspapers under the couch in the wastebasket.
  (@embed_qct) *EmQ -GP:* He asked her if she had decided to cram the old newspapers under the couch in the guestroom.  
  
  (@embed_dgp) *EmD +GP:* She told him that she had decided to cram the old newspapers under the couch in the wastebasket.
  (@embed_dct) *EmD -GP:* She told him that she had decided to cram the old newspapers under the couch in the guestroom.

The prosody of an embedded question, as in (@embed_qgp), does not differ from that of a sentence like (@embed_dgp); but, the semantic properties and some of the pragmatic properties of the embedded clause in (@embed_qgp) are the same as in one of the Q+GP sentences from the study reported here.

  (@qgp_re) Had she decided to crammed the old newspapers under the couch in the wastebasket?

These stimuli could be used in another reading aloud study, like the one just reported, or could be used with ERP or eye-tracking methodology. If a behavioral correlate of the intuition is found with embedded question stimuli, that would be strong evidence that the explanation is semantic or pragmatic rather than prosodic.

### Event-related potential (ERP) {#erp}

An event-related potential study of the phenomenon could provide a number of useful insights. It has been shown that the so-called closure-positive shift (CPS), an ERP component that is elicited by the presence of a prosodic break or a comma, is present even in silent reading. This has a number of important implications for research on the phenomena at hand. 

For one, it provides a way to distinguish between a hesitation after PP1 and a true prosodic break; which is to say, a way to distingiush between an unencumbered reading with a prosodic break after PP1 and a garden-path effect that, in the study reported here, resulted in a hesitation at that point. An ERP measure time-locked to that position should show a CPS component when the reader succesfully incoprorates PP1 as a modifier of the object and then changes branching direction to incorporate PP2 as the goal argument; conversely, a reader who is experiencing a garden-path effect at that point should instead or additionally show a P600 component (@p600addscps found that the two components can be additive). If the participant were concurrently recorded reading the sentence aloud, definitive data could be created which could then be used to either train a linguist or perhaps a neural network to distinguish between hesitations (P600s) and prosodic breaks (CPSes). The trained model or linguist could then return to the recordings collect for this study and make definitive judgements regarding the nature of the breaks reported here, and clarify the results of this study.

It would also be useful to see if repeated readings of the same garden-path sentence reduces the P600; if so, is the rate of decay faster for interrogatives than declaratives? Such a phenomenon could be a correlate of the reported intuition[^qerp].

[^qerp]: QUESTION: Do P600 components reflect the strength of a garden path? Also, do I need to go into any detail here about how ERP works or what the properties of P600 and CPS are?


