---
title: "R1 Delay category (median)"
author: "Tyler J. Peckenpaugh"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: no
    latex_engine: xelatex
fontsize: 12pt
mainfont: Georgia
---

|  **** | **Reading 1** |  |  |  | **Reading 2** |  |  |  |
| :--- | :---: | --- | --- | --- | :---: | --- | --- | --- |
|  **** | **Mod** |  | **Arg** |  | **Mod** |  | **Arg** |  |
|  **** | **D** | **Q** | **D** | **Q** | **D** | **Q** | **D** | **Q** |
|  Both | 54.1 | 43.0 | 72.1 | 71.7 | 54.1 | 43.0 | 72.1 | 71.7 |
|  OBJ only | 31.1 | 31.4 | 0.8 | 2.5 | 31.1 | 31.4 | 0.8 | 2.5 |
|  PP1 only | 14.8 | 25.6 | 27.0 | 25.8 | 14.8 | 25.6 | 27.0 | 25.8 |

```{r results_setup,echo=F,warning=F,message=F}
knitr::opts_chunk$set(
  echo=FALSE, 
  warning=FALSE, 
  message=FALSE, 
  fig.height = 3, 
  fig.pos="H"
)
options(tinytex.verbose = TRUE)
bgcolor <- "#ffffff"
par(bg = bgcolor)

kable <- function(data,...) {
  knitr::kable(data, booktabs = TRUE,...)
}

ms2s<-function(x,d=2) {
  return (format(x/1000,nsmall=d,digits=d))
}

library(extrafont)
library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(psych)

hidel <- 1050
lodel <- 901
```

# Reading 1 delay category

```{r delcat}

mdata <- read_csv("../drafts/export/prosody_data.csv")
raw_rs_file <- read_csv("../csvs/merged.csv")
raw_rs_file$UID <- as.factor(with(raw_rs_file, paste(Participant,Item,Reading,sep="-")))
raw_rs_file <- subset(raw_rs_file,UID %in% mdata$UID)
raw_rs <- subset(raw_rs_file,!isFiller)
raw_rs$reading <- raw_rs$Reading


r1data <- subset(raw_rs, reading == 1)
r1delByP <- aggregate(r1data$Leading,by=list("Participant" = r1data$Participant), FUN=median)

r1delByP$r1DelCat <- ifelse(r1delByP$x < lodel, "FAST", ifelse(
  r1delByP$x > hidel, "SLOW", "NORMAL"
))
catdes <- sprintf("FAST median R1 delay $\\\\leq$ %0.2fs. SLOW median R1 delay $>$ %0.2fs",lodel/1000,hidel/1000)

tab <- table(r1delByP$r1DelCat)
tab[c(1,3)] %>% 
  kable(col.names=c("Delay category", "n"), caption="Participants by Reading 1 delay category") %>% 
  kable_styling(full_width = F,latex_options = c("hold_position")) %>% 
  footnote(catdes,escape = F)

colnames(r1delByP)[2] <- "r1delbyp"

mdata$Participant <- mdata$SID
mdata<-merge(r1delByP,mdata,by="Participant")
mdata <- subset(mdata,simple2lvl != "NEITHER")
mdata$condition <- factor(mdata$condition,levels(factor(mdata$condition))[c(1,3,2,4)])
r1deldes <- describe(r1delByP$r1delbyp)
```

Reading 1 (R1) delay is the amount of time between the initial display of a sentence and the start of phonation. Participants' median R1 delay ranged from `r ms2s(r1deldes$min)`s to `r ms2s(r1deldes$max)`s with a standard deviation of `r ms2s(r1deldes$sd)`s. As a way of analyzing the protocol, and the extent to which participants performed as expected, participants were categorized based on their median R1 delay. In what follows, a fast median R1 delay is shorter than or equal to `r ms2s(lodel,2)`s, and a slow one is longer than `r ms2s(hidel,2)`s, resulting in 12 participants per category. Ten other participants categorized as "normal" and ignored. The distribution of participants across categories is shown in table \ref{tab:delcat}. These calculations were done over Reading 1 of experimental items (n = `r nrow(r1data)`, with `r 16*32 - nrow(r1data)` missing items).

```{r facet, fig.cap="Plot of pattern proportions per condition",fig.height=4,fig.pos="H"}
adata<-mdata
library('ggthemes')
if(!"reading" %in% names(adata)){
  adata$reading <- adata$Reading
}
loadfonts(device="pdf")
adata$Reading <- paste("Reading", adata$reading)

pdata <- subset(adata, r1DelCat != "NORMAL")

ggplot(pdata,    
    aes(
      x=condition,
      y=..count..,
      fill=reorder(
        simple2lvl,
        ifelse(simple2lvl=="PP1", 1, ifelse(simple2lvl=="OBJ",3,2))
      )
    )) +
  geom_bar(position="fill",color="black",width=0.5) +
  labs(fill="Pattern",x=" ",y=" ",caption="FAST n=12, SLOW n = 12") +
  facet_grid(rows = vars(Reading),cols=vars(r1DelCat)) +
  scale_y_continuous(labels=scales::percent) + 
  scale_fill_brewer(breaks=c("PP1", "BOTH", "OBJ"), palette="Greys") +
  theme_minimal(base_size = 12, base_family="Georgia")
```
<!--
```{r both}
adata<-mdata

fast <- subset(mdata,r1DelCat == "FAST")
slow <- subset(mdata,r1DelCat == "SLOW")

fastTabN <- xtabs(~simple2lvl + condition, data=fast) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)  
slowTabN <- xtabs(~simple2lvl + condition, data=slow) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)

fastTab <- xtabs(~simple2lvl + condition, data=fast) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTab <- xtabs(~simple2lvl + condition, data=slow) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)


tab<-cbind(fastTab,slowTab[2:5])
ntab<-cbind(fastTabN,slowTabN[2:5])
tab[2:9]  <- format(tab[2:9] * 100,nsmall=1,digits=1)
ntab[2:9] <- format(ntab[2:9],digits=0)

tab$m  <- "%"
ntab$m <- "n"
tab <- rbind(tab,ntab)

colnames(tab)[1] <- "Pattern"
colnames(tab)[6:9] <- paste0(colnames(tab)[6:9],".slow")
tab <- arrange(tab,Pattern)
tab <- tab %>% select(Pattern,m,everything())


tab %>% kable(
  align="c",
  caption="Simple break pattern by condition and R1 delay category, both readings",
  col.names=c(" "," ", rep(c("D -GP","Q -GP", "D +GP", "Q +GP"),2)), 
  digits=1
) %>% kable_styling() %>%collapse_rows(columns = 1) %>%
  add_header_above(c(" " = 2,"FAST"=4,"SLOW"=4)) 

```
-->


```{r rsplit}
sdata<-subset(mdata,Reading==1)

fast <- subset(sdata,r1DelCat == "FAST")
slow <- subset(sdata,r1DelCat == "SLOW")

fastTabN <- xtabs(~simple2lvl + condition, data=fast) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)  
slowTabN <- xtabs(~simple2lvl + condition, data=slow) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)


fastTab <- xtabs(~simple2lvl + condition, data=fast) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTab <- xtabs(~simple2lvl + condition, data=slow) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)


tab<-cbind(fastTab,slowTab[2:5])
ntab<-cbind(fastTabN,slowTabN[2:5])
tab[2:9]  <- format(tab[2:9] * 100,nsmall=1,digits=1)
ntab[2:9] <- format(ntab[2:9],digits=0)


tab$m  <- "%"
ntab$m <- "n"
tab <- rbind(tab,ntab)

colnames(tab)[1] <- "Pattern"
colnames(tab)[6:9] <- paste0(colnames(tab)[6:9],".slow")
tab <- arrange(tab,Pattern)
tab <- tab %>% select(Pattern,m,everything())

r1tab<-tab


sdata<-subset(mdata,Reading==2)

fast <- subset(sdata,r1DelCat == "FAST")
slow <- subset(sdata,r1DelCat == "SLOW")

fastTabN <- xtabs(~simple2lvl + condition, data=fast) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)  
slowTabN <- xtabs(~simple2lvl + condition, data=slow) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)

fastTab <- xtabs(~simple2lvl + condition, data=fast) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTab <- xtabs(~simple2lvl + condition, data=slow) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)


tab<-cbind(fastTab,slowTab[2:5])
ntab<-cbind(fastTabN,slowTabN[2:5])
tab[2:9]  <- format(tab[2:9] * 100,nsmall=1,digits=1)
ntab[2:9] <- format(ntab[2:9],digits=0)


tab$m  <- "%"
ntab$m <- "n"
tab <- rbind(tab,ntab)

colnames(tab)[1] <- "Pattern"
colnames(tab)[6:9] <- paste0(colnames(tab)[6:9],".slow")
tab <- arrange(tab,Pattern)
tab <- tab %>% select(Pattern,m,everything())


tab <- rbind(r1tab,tab)

tab %>% kable(
  align="c",
  caption="Simple break pattern by condition and R1 delay category",
  col.names=c(" "," ", rep(c("D -GP","Q -GP", "D +GP", "Q +GP"),2)), 
  digits=1
) %>% kable_styling() %>%
  column_spec(6,border_right = T) %>%
  collapse_rows(columns = 1) %>%
  add_header_above(c(" " = 2,"FAST (n=12)"=4,"SLOW (n=12)"=4)) %>%
  pack_rows(index = c("Reading 1" = 6, "Reading 2" = 6))%>% 
  footnote(catdes,escape = F)


```

Table \ref{tab:rsplit} shows the number of occurrences and percentage per condition for each readings. Participants' median R1 delay can be seen in Appendix \ref{sec:r1dbyp}.