---
title: "R1 Delay category"
author: "Tyler J. Peckenpaugh"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: no
    latex_engine: xelatex
fontsize: 12pt
mainfont: Georgia
geometry: "left=2cm,right=2cm,top=3cm,bottom=3cm"
---
```{r results_setup,echo=F,warning=F,message=F}
knitr::opts_chunk$set(
  echo=FALSE, 
  warning=FALSE, 
  message=FALSE, 
  fig.height = 3, 
  fig.pos="H"
)
options(tinytex.verbose = TRUE)
bgcolor <- "#ffffff"
par(bg = bgcolor)

kable <- function(data,...) {
  knitr::kable(data, booktabs = TRUE,...)
}

library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)

mdata <- read_csv("../drafts/export/prosody_data.csv")

```
# Participant categories of reading 1 delay

A brief analysis of the effect of Reading 1 delay on the prosodic patterns produced is presented.

```{r delcat}
raw_rs_file <- read_csv("../csvs/merged.csv")
raw_rs_file$UID <- as.factor(with(raw_rs_file, paste(Participant,Item,Reading,sep="-")))
raw_rs_file <- subset(raw_rs_file,UID %in% mdata$UID)
raw_rs <- subset(raw_rs_file,!isFiller)
raw_rs$reading <- raw_rs$Reading
raw_rs$Reading <- ifelse(raw_rs$reading == 1, "Cold", "Previewed")

r1data <- subset(raw_rs, reading == 1)
r1delByP <- aggregate(r1data$Leading,by=list("Participant" = r1data$Participant), FUN=mean)
hidel <- 1200
lodel <- 900
r1delByP$r1DelCat <- ifelse(r1delByP$x < lodel, "FAST", ifelse(
  r1delByP$x > hidel, "SLOW", "NORMAL"
))
catdes <- sprintf("FAST < %0.1fs; SLOW > %0.1fs",lodel/1000,hidel/1000)

table(r1delByP$r1DelCat) %>% 
  kable(col.names=c("Delay category", "n"), caption="Participants by Reading 1 delay category") %>% 
  kable_styling(full_width = F) %>% 
  footnote(catdes)

colnames(r1delByP)[2] <- "r1delbyp"

mdata$Participant <- mdata$SID
mdata<-merge(r1delByP,mdata,by="Participant")
mdata <- subset(mdata,simple2lvl != "NEITHER")

```

First, participants were categorized based on their mean Reading 1 delay, with a fast mean Reading 1 delay being less than `r round(lodel/1000,2)`s, and a slow one being more than `r round(hidel/1000,2)`s. Everything else is categorized as "normal" and ignored. The distribution of participants across categories is shown in table \ref{tab:delcat}. Each participant's mean is shown in table \ref{tab:r1delByP}

```{r}
fast <- subset(mdata,r1DelCat == "FAST")
slow <- subset(mdata,r1DelCat == "SLOW")
fastTabN <- xtabs(~simple2lvl + condition, data=fast) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTabN <- xtabs(~simple2lvl + condition, data=slow) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
fastTab <- xtabs(~simple2lvl + condition, data=fast) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTab <- xtabs(~simple2lvl + condition, data=slow) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)

tab<-cbind(fastTab,slowTab[2:5])
ntab<-cbind(fastTabN,slowTabN[2:5])
tab[2:9] <- tab[2:9] * 100
tab$m  <- "%"
ntab$m <- "n"
tab <- rbind(tab,ntab)


colnames(tab)[1] <- "Pattern"
colnames(tab)[6:9] <- paste0(colnames(tab)[6:9],".slow")
tab <- arrange(tab,Pattern)
tab <- tab %>% select(Pattern,m,everything())


tab %>% kable(
  caption="Simple break pattern by condition and R1 delay category, both readings",
  col.names=c(" ","m","D -GP", "D +GP", "Q -GP", "Q +GP", "D -GP", "D +GP", "Q -GP", "Q +GP"), 
  digits=2
) %>% kable_styling() %>%collapse_rows(columns = 1) %>%
  add_header_above(c(" " = 2,"FAST"=4,"SLOW"=4)) 

```
```{r}
cheatdata <- subset(mdata,r1DelCat == "FAST" & Reading == 2)
pp1tab<-xtabs(PP1 ~ Condition_GP + Condition_Q, data=cheatdata)
objtab<-xtabs(OBJ ~ Condition_GP + Condition_Q, data=cheatdata)
tot<-as.data.frame(xtabs( ~ Condition_GP + Condition_Q, data=cheatdata))


pp1<-as.data.frame(pp1tab)
pp1$tot<-tot$Freq
pp1$pct <- pp1$Freq/pp1$tot
pp1$pretty <- paste0(
  pp1$Freq, 
  " (", 
  sprintf("%.1f",pp1$pct * 100), 
  "%)"
)
pp1disp <- pp1 %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

obj<-as.data.frame(objtab)
obj$tot<-tot$Freq
obj$pct <- obj$Freq/obj$tot
obj$pretty <- paste0(
  obj$Freq, 
  " (", 
  sprintf("%.1f",obj$pct * 100), 
  "%)"
)
objdisp <- obj %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

pp1obj <- cbind(pp1disp,objdisp[2:3])
pp1obj[[1]] <- c("+GP","-GP")
fast <- pp1obj
```

```{r}
cheatdata <- subset(mdata,r1DelCat == "SLOW" & Reading == 2)
pp1tab<-xtabs(PP1 ~ Condition_GP + Condition_Q, data=cheatdata)
objtab<-xtabs(OBJ ~ Condition_GP + Condition_Q, data=cheatdata)
tot<-as.data.frame(xtabs( ~ Condition_GP + Condition_Q, data=cheatdata))


pp1<-as.data.frame(pp1tab)
pp1$tot<-tot$Freq
pp1$pct <- pp1$Freq/pp1$tot
pp1$pretty <- paste0(
  pp1$Freq, 
  " (", 
  sprintf("%.1f",pp1$pct * 100), 
  "%)"
)
pp1disp <- pp1 %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

obj<-as.data.frame(objtab)
obj$tot<-tot$Freq
obj$pct <- obj$Freq/obj$tot
obj$pretty <- paste0(
  obj$Freq, 
  " (", 
  sprintf("%.1f",obj$pct * 100), 
  "%)"
)
objdisp <- obj %>%
  subset(select=c(
    "Condition_GP",
    "Condition_Q",
    "pretty"
  )) %>% 
  spread(Condition_Q,pretty)

pp1obj <- cbind(pp1disp,objdisp[2:3])
pp1obj[[1]] <- c("+GP","-GP")
```

The break patterns for slow and fast participants are shown in table \ref{tab:cheaterBreaks}.

```{r cheaterBreaks}
cat <-cbind(fast, pp1obj[2:5])
kable(
  cat,
  caption="Percentage of R2 recordings with PP1/OBJ break by condition and R1 delay category",
  col.names = c("", "D", "Q","D", "Q","D", "Q","D", "Q")
) %>% 
  add_header_above(c(" "=1,"PP1 Break"=2,"OBJ Break"=2,"PP1 Break"=2,"OBJ Break"=2)) %>%
  add_header_above(c(" "=1,"Fast R1 delay"=4,"Slow R1 delay"=4)) %>%
  kable_styling(latex_options = c("hold_position","scale_down"))
```

```{r}
cheatdata <- subset(mdata,r1DelCat == "FAST" & Reading == 2)
twolv<-xtabs( ~ two_level_prosody + condition, data=cheatdata) %>%
  prop.table(margin=2)

twolvFAST <- twolv * 100

cheatdata <- subset(mdata,r1DelCat == "SLOW" & Reading == 2)
twolv<-xtabs( ~ two_level_prosody + condition, data=cheatdata) %>%
  prop.table(margin=2)

twolv <- cbind(twolvFAST, twolv * 100)

twolv %>% 
  kable(
    caption="Prosodic pattern by condition, SLOW/FAST R1 delay category (Reading 2)",
    digits = 2
  ) %>%
  add_header_above(c(" " = 1, "FAST" = 4, "SLOW" = 4)) %>%
  kable_styling(latex_options = c("hold_position"))
```
```{r}
cheatdata <- subset(mdata,r1DelCat == "FAST" & Reading == 1)
twolv<-xtabs( ~ two_level_prosody + condition, data=cheatdata) %>%
  prop.table(margin=2)

twolvFAST <- twolv * 100

cheatdata <- subset(mdata,r1DelCat == "SLOW" & Reading == 1)
twolv<-xtabs( ~ two_level_prosody + condition, data=cheatdata) %>%
  prop.table(margin=2)

twolv <- cbind(twolvFAST, twolv * 100)

twolv %>% 
  kable(
    caption="Prosodic pattern by condition, SLOW/FAST R1 delay category (Reading 1)",
    digits = 2
  ) %>%
  add_header_above(c(" " = 1, "FAST" = 4, "SLOW" = 4)) %>%
  kable_styling(latex_options = c("hold_position"))
```

```{r r1delByP}
df <- r1delByP
df$r1delbyp <- df$r1delbyp / 1000

df %>%  
  arrange(df$r1delbyp) %>% 
  kable(
    col.names = c("Participant", "Mean R1 delay","R1 delay category"), 
    caption="Mean reading 1 delay (seconds) and R1 delay category by participant", 
    digits = 2
  ) %>% kable_styling(latex_options = c("repeat_header"))%>% 
  footnote(catdes)
```