---
title: "R1 Delay category"
author: "Tyler J. Peckenpaugh"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: no
    latex_engine: xelatex
fontsize: 12pt
mainfont: Georgia
---
```{r results_setup,echo=F,warning=F,message=F}
knitr::opts_chunk$set(
  echo=FALSE, 
  warning=FALSE, 
  message=FALSE, 
  fig.height = 3, 
  fig.pos="H"
)
options(tinytex.verbose = TRUE)
bgcolor <- "#ffffff"
par(bg = bgcolor)

kable <- function(data,...) {
  knitr::kable(data, booktabs = TRUE,...)
}

library(readr)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(ggplot2)
hidel <- 1200
lodel <- 900
```

# Reading 1 delay category

First, participants were categorized based on their mean Reading 1 delay, with a fast mean Reading 1 delay being less than `r round(lodel/1000,2)`s, and a slow one being more than `r round(hidel/1000,2)`s. Everything else is categorized as "normal" and ignored. The distribution of participants across categories is shown in table \ref{tab:delcat}. 

```{r delcat}

mdata <- read_csv("../drafts/export/prosody_data.csv")
raw_rs_file <- read_csv("../csvs/merged.csv")
raw_rs_file$UID <- as.factor(with(raw_rs_file, paste(Participant,Item,Reading,sep="-")))
raw_rs_file <- subset(raw_rs_file,UID %in% mdata$UID)
raw_rs <- subset(raw_rs_file,!isFiller)
raw_rs$reading <- raw_rs$Reading
raw_rs$Reading <- ifelse(raw_rs$reading == 1, "Cold", "Previewed")

r1data <- subset(raw_rs, reading == 1)
r1delByP <- aggregate(r1data$Leading,by=list("Participant" = r1data$Participant), FUN=mean)

r1delByP$r1DelCat <- ifelse(r1delByP$x < lodel, "FAST", ifelse(
  r1delByP$x > hidel, "SLOW", "NORMAL"
))
catdes <- sprintf("FAST < %0.1fs; SLOW > %0.1fs",lodel/1000,hidel/1000)

table(r1delByP$r1DelCat) %>% 
  kable(col.names=c("Delay category", "n"), caption="Participants by Reading 1 delay category") %>% 
  kable_styling(full_width = F,latex_options = c("hold_position")) %>% 
  footnote(catdes)

colnames(r1delByP)[2] <- "r1delbyp"

mdata$Participant <- mdata$SID
mdata<-merge(r1delByP,mdata,by="Participant")
mdata <- subset(mdata,simple2lvl != "NEITHER")
mdata$condition <- factor(mdata$condition,levels(factor(mdata$condition))[c(1,3,2,4)])
```

```{r facet, fig.cap="Plot of pattern proportions per condition"}
adata<-mdata
library('ggthemes')
if(!"reading" %in% names(adata)){
  adata$reading <- adata$Reading
}
adata$Reading <- paste("Reading", adata$reading)

pdata <- subset(adata, r1DelCat != "NORMAL")

ggplot(pdata,    
    aes(
      x=condition,
      y=..count..,
      fill=reorder(
        simple2lvl,
        ifelse(simple2lvl=="PP1", 1, ifelse(simple2lvl=="OBJ",3,2))
      )
    )) +
  geom_bar(position="fill",color="black",width=0.5) +
  labs(fill="Pattern",x=" ",y=" ") +
  facet_grid(rows = vars(Reading),cols=vars(r1DelCat)) +
  scale_y_continuous(labels=scales::percent) + 
  scale_fill_brewer(breaks=c("PP1", "BOTH", "OBJ"), palette="Greys") +
  theme_minimal(base_size = 12)
```

```{r both}
adata<-mdata

fast <- subset(mdata,r1DelCat == "FAST")
slow <- subset(mdata,r1DelCat == "SLOW")
fastTabN <- xtabs(~simple2lvl + condition, data=fast) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTabN <- xtabs(~simple2lvl + condition, data=slow) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
fastTab <- xtabs(~simple2lvl + condition, data=fast) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTab <- xtabs(~simple2lvl + condition, data=slow) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)

tab<-cbind(fastTab,slowTab[2:5])
ntab<-cbind(fastTabN,slowTabN[2:5])
tab[2:9] <- tab[2:9] * 100
tab$m  <- "%"
ntab$m <- "n"
tab <- rbind(tab,ntab)

colnames(tab)[1] <- "Pattern"
colnames(tab)[6:9] <- paste0(colnames(tab)[6:9],".slow")
tab <- arrange(tab,Pattern)
tab <- tab %>% select(Pattern,m,everything())


tab %>% kable(
  caption="Simple break pattern by condition and R1 delay category, both readings",
  col.names=c(" ","m","D -GP", "D +GP", "Q -GP", "Q +GP", "D -GP", "D +GP", "Q -GP", "Q +GP"), 
  digits=1
) %>% kable_styling() %>%collapse_rows(columns = 1) %>%
  add_header_above(c(" " = 2,"FAST"=4,"SLOW"=4)) 

```


```{r r1}
mdata<-subset(adata,Reading==1)
fast <- subset(mdata,r1DelCat == "FAST")
slow <- subset(mdata,r1DelCat == "SLOW")
fastTabN <- xtabs(~simple2lvl + condition, data=fast) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTabN <- xtabs(~simple2lvl + condition, data=slow) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
fastTab <- xtabs(~simple2lvl + condition, data=fast) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTab <- xtabs(~simple2lvl + condition, data=slow) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)

tab<-cbind(fastTab,slowTab[2:5])
ntab<-cbind(fastTabN,slowTabN[2:5])
tab[2:9] <- tab[2:9] * 100
tab$m  <- "%"
ntab$m <- "n"
tab <- rbind(tab,ntab)

colnames(tab)[1] <- "Pattern"
colnames(tab)[6:9] <- paste0(colnames(tab)[6:9],".slow")
tab <- arrange(tab,Pattern)
tab <- tab %>% select(Pattern,m,everything())


tab %>% kable(
  caption="Simple break pattern by condition and R1 delay category,  Reading 1",
  col.names=c(" ","m","D -GP", "D +GP", "Q -GP", "Q +GP", "D -GP", "D +GP", "Q -GP", "Q +GP"), 
  digits=1
) %>% kable_styling() %>%collapse_rows(columns = 1) %>%
  add_header_above(c(" " = 2,"FAST"=4,"SLOW"=4)) 

```


```{r r2}
mdata<-subset(adata,Reading==2)
fast <- subset(mdata,r1DelCat == "FAST")
slow <- subset(mdata,r1DelCat == "SLOW")
fastTabN <- xtabs(~simple2lvl + condition, data=fast) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTabN <- xtabs(~simple2lvl + condition, data=slow) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
fastTab <- xtabs(~simple2lvl + condition, data=fast) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)
slowTab <- xtabs(~simple2lvl + condition, data=slow) %>% 
  prop.table(margin=2) %>% 
  as.data.frame() %>% 
  spread(condition, Freq)

tab<-cbind(fastTab,slowTab[2:5])
ntab<-cbind(fastTabN,slowTabN[2:5])
tab[2:9] <- tab[2:9] * 100
tab$m  <- "%"
ntab$m <- "n"
tab <- rbind(tab,ntab)

colnames(tab)[1] <- "Pattern"
colnames(tab)[6:9] <- paste0(colnames(tab)[6:9],".slow")
tab <- arrange(tab,Pattern)
tab <- tab %>% select(Pattern,m,everything())


tab %>% kable(
  caption="Simple break pattern by condition and R1 delay category,  Reading 2",
  col.names=c(" ","m","D -GP", "D +GP", "Q -GP", "Q +GP", "D -GP", "D +GP", "Q -GP", "Q +GP"), 
  digits=1
) %>% kable_styling() %>%collapse_rows(columns = 1) %>%
  add_header_above(c(" " = 2,"FAST"=4,"SLOW"=4)) 

```


Table \ref{tab:both} shows the number of occurrences and percentage per condition for both readings. Table  \ref{tab:r1} shows the same for just reading 1, and table \ref{tab:r2} for just reading 2.